\chapter{Present Simple and Present Continuous}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item How to form and use the Present Simple tense
    \item How to form and use the Present Continuous tense
    \item The difference between routines (Present Simple) and actions happening now (Present Continuous)
    \item Stative verbs that don't use continuous forms
    \item Time expressions for each tense
\end{itemize}

\section{Reading Context}
\begin{readingbox}[title=Dialogue: A Typical Day vs. Right Now]
\textbf{Emma:} Hi Tom! What are you doing?\\
\textbf{Tom:} I\textbf{'m reading} a book about business. What about you?\\
\textbf{Emma:} I usually \textbf{work} in the office, but today I\textbf{'m working} from home.\\
\textbf{Tom:} That's nice. Do you \textbf{like} working from home?\\
\textbf{Emma:} Yes, I do. I \textbf{prefer} it because I \textbf{don't spend} time commuting.\\
\textbf{Tom:} I \textbf{understand}. The kids \textbf{are playing} outside right now, so it's quiet.\\
\textbf{Emma:} Perfect. By the way, what time do you usually \textbf{have} lunch?\\
\textbf{Tom:} I normally \textbf{eat} at 1 PM, but today I\textbf{'m having} lunch early.
\end{readingbox}

\section{Grammar Focus: Present Simple}

The Present Simple describes \textbf{routines, habits, facts, and general truths}.

\begin{grammarbox}[title=Present Simple Structure]
\textbf{Affirmative:}
\begin{itemize}
    \item I/You/We/They + \textbf{verb} (base form)
    \item He/She/It + \textbf{verb + s/es}
\end{itemize}

\textbf{Negative:}
\begin{itemize}
    \item I/You/We/They + \textbf{don't} + verb
    \item He/She/It + \textbf{doesn't} + verb
\end{itemize}

\textbf{Questions:}
\begin{itemize}
    \item \textbf{Do} + I/you/we/they + verb?
    \item \textbf{Does} + he/she/it + verb?
\end{itemize}

\textbf{Examples:}
\begin{itemize}
    \item I \textbf{work} in an office. \trans{Trabajo en una oficina}
    \item She \textbf{works} every day. \trans{Ella trabaja todos los días}
    \item They \textbf{don't like} coffee. \trans{No les gusta el café}
    \item \textbf{Does} he \textbf{speak} English? \trans{¿Habla inglés?}
\end{itemize}
\end{grammarbox}

\subsection{Spelling Rules for Third Person (he/she/it)}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Rule} & \textbf{Example} & \textbf{Result} \\
\hline
Most verbs: add -s & work, play & works, plays \\
\hline
Ends in -s, -sh, -ch, -x, -o: add -es & wash, go, miss & washes, goes, misses \\
\hline
Ends in consonant + y: change to -ies & study, carry & studies, carries \\
\hline
Ends in vowel + y: add -s & play, enjoy & plays, enjoys \\
\hline
\end{tabular}
\caption{Third person singular spelling rules}
\end{table}

\section{Grammar Focus: Present Continuous}

The Present Continuous describes \textbf{actions happening now} or \textbf{temporary situations}.

\begin{grammarbox}[title=Present Continuous Structure]
\textbf{Structure:} Subject + \textbf{am/is/are} + \textbf{verb-ing}

\textbf{Affirmative:}
\begin{itemize}
    \item I \textbf{am working} (I'm working)
    \item He/She/It \textbf{is working} (He's working)
    \item We/You/They \textbf{are working} (They're working)
\end{itemize}

\textbf{Negative:}
\begin{itemize}
    \item I \textbf{am not} working (I'm not working)
    \item He \textbf{is not} working (He isn't working)
    \item They \textbf{are not} working (They aren't working)
\end{itemize}

\textbf{Questions:}
\begin{itemize}
    \item \textbf{Am} I working?
    \item \textbf{Is} he/she/it working?
    \item \textbf{Are} you/we/they working?
\end{itemize}

\textbf{Examples:}
\begin{itemize}
    \item I \textbf{am studying} right now. \trans{Estoy estudiando ahora}
    \item She \textbf{is reading} a book. \trans{Ella está leyendo un libro}
    \item They \textbf{aren't listening}. \trans{No están escuchando}
\end{itemize}
\end{grammarbox}

\subsection{Spelling Rules for -ING Forms}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Rule} & \textbf{Example} & \textbf{Result} \\
\hline
Most verbs: add -ing & work, play & working, playing \\
\hline
Ends in -e: remove e, add -ing & make, write & making, writing \\
\hline
One syllable CVC: double consonant & sit, run, stop & sitting, running, stopping \\
\hline
Two syllables (stressed): double & begin, prefer & beginning, preferring \\
\hline
Ends in -ie: change to -ying & lie, die & lying, dying \\
\hline
\end{tabular}
\caption{Verb + -ing spelling rules}
\end{table}

\section{Key Differences: Simple vs. Continuous}

\begin{table}[h]
\centering
\begin{tabular}{|p{6cm}|p{6cm}|}
\hline
\textbf{Present Simple} & \textbf{Present Continuous} \\
\hline
Routines and habits & Actions happening now \\
\hline
I \textbf{go} to work by bus every day. & I \textbf{am going} to work by bus right now. \\
\hline
General truths and facts & Temporary situations \\
\hline
She \textbf{lives} in London. & She \textbf{is living} in a hotel this week. \\
\hline
Permanent situations & Actions in progress \\
\hline
He \textbf{works} in a bank. & He \textbf{is working} on a project. \\
\hline
\end{tabular}
\caption{Comparing Present Simple and Continuous}
\end{table}

\section{Stative Verbs (No Continuous)}

Some verbs describe \textbf{states} (not actions) and are NOT normally used in continuous forms.

\begin{vocabbox}[title=Common Stative Verbs]
\begin{itemize}
    \item \textbf{Feelings:} like, love, hate, prefer, want, need
    \item \textbf{Thinking:} know, understand, believe, think (opinion), remember
    \item \textbf{Senses:} see, hear, smell, taste (involuntary)
    \item \textbf{Possession:} have (possess), own, belong
    \item \textbf{Being:} be, seem, appear
\end{itemize}

\textbf{Examples:}
\begin{itemize}
    \item I \textbf{like} pizza. \checkmark (NOT: I'm liking pizza. \textbf{X})
    \item She \textbf{knows} the answer. \checkmark (NOT: She's knowing. \textbf{X})
    \item We \textbf{have} a car. \checkmark (possession)
    \item We \textbf{are having} dinner. \checkmark (action - eating)
\end{itemize}
\end{vocabbox}

\section{Time Expressions}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Present Simple} & \textbf{Present Continuous} \\
\hline
always, usually, often & now, right now, at the moment \\
sometimes, rarely, never & today, this week/month \\
every day/week/month & currently, at present \\
on Mondays, in the morning & Look!, Listen! \\
\hline
\end{tabular}
\caption{Common time expressions}
\end{table}

\section{Practice Exercises}

\subsection{Exercise 1: Present Simple - Fill in the Blanks}
Complete with the correct form of the verb in parentheses.

\begin{enumerate}
    \item She \underline{\hspace{3cm}} (work) in a hospital.
    \item They \underline{\hspace{3cm}} (not/like) spicy food.
    \item \underline{\hspace{3cm}} you \underline{\hspace{3cm}} (speak) French?
    \item He \underline{\hspace{3cm}} (go) to the gym every morning.
    \item We \underline{\hspace{3cm}} (not/watch) TV in the evening.
\end{enumerate}

\subsection{Exercise 2: Present Continuous - Fill in the Blanks}
Complete with the correct form of the verb in parentheses.

\begin{enumerate}
    \item I \underline{\hspace{3cm}} (read) a book right now.
    \item She \underline{\hspace{3cm}} (not/work) today.
    \item \underline{\hspace{3cm}} they \underline{\hspace{3cm}} (come) to the party?
    \item We \underline{\hspace{3cm}} (have) lunch at the moment.
    \item He \underline{\hspace{3cm}} (study) for his exam.
\end{enumerate}

\subsection{Exercise 3: Simple or Continuous?}
Choose the correct tense.

\begin{enumerate}
    \item I (work / am working) in London. I started this job 5 years ago.
    \item Look! It (rains / is raining).
    \item She (doesn't like / isn't liking) coffee. She prefers tea.
    \item What (do you do / are you doing)? I'm a teacher.
    \item What (do you do / are you doing) right now? I'm reading.
    \item I (think / am thinking) he's wrong. (opinion)
    \item I (think / am thinking) about my vacation. (process)
\end{enumerate}

\subsection{Exercise 4: Correct the Errors}
Find and correct the mistakes.

\begin{enumerate}
    \item She is liking chocolate very much.
    \item He go to school every day.
    \item Are you knowing the answer?
    \item I'm not understanding this lesson.
    \item They doesn't work on Sundays.
\end{enumerate}

\subsection{Exercise 5: Writing Task}
Write two paragraphs (60-80 words each):
\begin{enumerate}
    \item Describe your typical day (use Present Simple)
    \item Describe what you are doing today (use Present Continuous)
\end{enumerate}

\begin{tcolorbox}[colback=white,height=8cm]
% Paragraph 1: My typical day...

\vspace{3cm}

% Paragraph 2: What I'm doing today...
\end{tcolorbox}

\section{Key Takeaways}
\begin{itemize}
    \item Use \textbf{Present Simple} for routines, habits, and facts
    \item Use \textbf{Present Continuous} for actions happening now or temporary situations
    \item Don't forget the \textbf{-s/-es} for third person in Present Simple
    \item \textbf{Stative verbs} (like, know, have) don't usually use continuous forms
    \item Pay attention to \textbf{time expressions} to choose the right tense
\end{itemize}
