\chapter{Writing Skills and Common Errors}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item How to plan, write, and check your written work
    \item Common errors in English writing and how to avoid them
    \item Linking words to connect ideas
    \item Common preposition collocations
\end{itemize}

\section{Reading Context}
\begin{readingbox}[title=Dialogue: Reviewing a Report]
\textbf{Editor:} I read your draft. It's good, but there are a few errors.\\
\textbf{Writer:} Really? I thought I checked it carefully.\\
\textbf{Editor:} You missed some articles. For example, you wrote "I have car" instead of "I have \textbf{a} car".\\
\textbf{Writer:} Oh, I see. I'm always worried \textbf{about} making mistakes with articles.\\
\textbf{Editor:} Don't worry. Also, use more linking words. \textbf{However}, the content is excellent.\\
\textbf{Writer:} Thanks. I'll give it another look. I'm interested \textbf{in} improving my writing.\\
\textbf{Editor:} Great. Remember to plan before you write next time.
\end{readingbox}

\section{Key Concepts: The Writing Process}

Good writing follows three key stages:

\begin{vocabbox}[title=The 3 Stages]
\begin{enumerate}
    \item \keyterm{Planning} (17\% of time): Brainstorm ideas, organize structure, make notes.
    \item \keyterm{Writing} (60\% of time): Write your first draft quickly, focus on content.
    \item \keyterm{Checking} (23\% of time): Proofread for errors, revise and improve.
\end{enumerate}
\end{vocabbox}

\section{Grammar Focus: Common Errors}

\begin{grammarbox}[title=Error Correction Guide]
\textbf{1. Missing Articles:}
\begin{itemize}
    \item Incorrect: I have car.
    \item Correct: I have \textbf{a} car.
\end{itemize}

\textbf{2. Wrong Prepositions:}
\begin{itemize}
    \item Incorrect: I'm interested for music.
    \item Correct: I'm interested \textbf{in} music.
\end{itemize}

\textbf{3. Run-on Sentences:}
\begin{itemize}
    \item Incorrect: I went home I was tired.
    \item Correct: I went home \textbf{because} I was tired.
\end{itemize}
\end{grammarbox}

\section{Vocabulary: Linking Words and Prepositions}

\begin{vocabbox}[title=Linking Words]
\begin{itemize}
    \item \textbf{Addition:} Furthermore, Moreover, Also \trans{Además}
    \item \textbf{Contrast:} However, Although, On the other hand \trans{Sin embargo}
    \item \textbf{Result:} Therefore, Consequently, As a result \trans{Por lo tanto}
    \item \textbf{Sequence:} First, Then, Next, Finally \trans{Primero, Luego...}
\end{itemize}
\end{vocabbox}

\begin{table}[h]
\centering
\begin{tabular}{|l|p{4cm}|p{5cm}|}
\hline
\textbf{Collocation} & \textbf{Spanish} & \textbf{Example} \\
\hline
Interested \textbf{in} & Interesado en & I'm interested \textbf{in} art. \\
\hline
Good \textbf{at} & Bueno en & She's good \textbf{at} math. \\
\hline
Responsible \textbf{for} & Responsable de & He's responsible \textbf{for} it. \\
\hline
Depend \textbf{on} & Depender de & It depends \textbf{on} you. \\
\hline
Afraid \textbf{of} & Miedo de & I'm afraid \textbf{of} spiders. \\
\hline
\end{tabular}
\caption{Common Preposition Collocations}
\end{table}

\section{Practice Exercises}

\subsection{Exercise 1: Find and Correct the Errors}
Each sentence has one error. Find and correct it.

\begin{enumerate}
    \item I have been living in London since 3 years.
    \item She is very good in cooking.
    \item I went to store and bought some milk.
    \item However I think we should wait.
\end{enumerate}

\subsection{Exercise 2: Complete with Linking Words}
Fill in the blanks with: \textit{However, Therefore, Furthermore, First}.

\begin{enumerate}
    \item I studied hard. \underline{\hspace{2cm}}, I passed the exam.
    \item The food was delicious. \underline{\hspace{2cm}}, it was expensive.
    \item \underline{\hspace{2cm}}, we need to buy tickets. Then, we can enter.
\end{enumerate}

\subsection{Exercise 3: Writing Practice}
Write a short paragraph (50-80 words) about your city using at least 2 linking words and 2 preposition collocations.

\begin{tcolorbox}[colback=white,height=5cm]
% Write your paragraph here
\end{tcolorbox}

\section{Key Takeaways}
\begin{itemize}
    \item Always plan before writing and leave time for checking.
    \item Watch out for missing articles and wrong prepositions.
    \item Use linking words (However, Therefore) to connect ideas.
    \item Memorize collocations: interested IN, good AT, depend ON.
\end{itemize}

\section{Online Practice}
\begin{tcolorbox}[colback=blue!5,colframe=blue!40!black,title=Resources to Practice]
Online PracticeReinforce what you have learned with these interactive exercises:
\begin{itemize}
    \item \textbf{Writing Skills:} \url{https://learnenglish.britishcouncil.org/skills/writing}
    \item \textbf{Linking Words:} \url{https://www.englishclub.com/writing/linking-words.htm}
    \item \textbf{Common Errors:} \url{https://www.perfect-english-grammar.com/common-mistakes.html}
    \item \textbf{Preposition Collocations:} \url{https://test-english.com/grammar-points/a2-b1/dependent-prepositions/}
\end{itemize}
\end{tcolorbox}