\chapter{Past Simple: Basic Past Tense}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item How to form the Past Simple with regular and irregular verbs
    \item When to use Past Simple
    \item How to make negative sentences and questions
    \item Common time expressions with Past Simple
    \item The difference between regular and irregular verbs
\end{itemize}

\section{Reading Context}
\begin{readingbox}[title=Dialogue: Last Weekend]
\textbf{Sarah:} What did you do last weekend?\\
\textbf{Mike:} I \textbf{went} to the cinema on Saturday. I \textbf{saw} the new James Bond movie.\\
\textbf{Sarah:} Oh! \textbf{Did} you \textbf{like} it?\\
\textbf{Mike:} Yes, I \textbf{loved} it! It \textbf{was} amazing. What about you?\\
\textbf{Sarah:} I \textbf{stayed} home. I \textbf{didn't go} out because I \textbf{had} a cold.\\
\textbf{Mike:} That's too bad. \textbf{Did} you \textbf{rest}?\\
\textbf{Sarah:} Yes. I \textbf{watched} TV and \textbf{read} a book. I also \textbf{talked} to my family on the phone.\\
\textbf{Mike:} Good. Are you feeling better now?\\
\textbf{Sarah:} Yes, much better, thanks!
\end{readingbox}

\section{Grammar Focus: Past Simple Basics}

The Past Simple describes \textbf{completed actions in the past}.

\begin{grammarbox}[title=Past Simple - When to Use]
Use Past Simple for:
\begin{itemize}
    \item \textbf{Finished actions:} I \textbf{visited} Rome last year. \trans{Visité Roma...}
    \item \textbf{Past habits:} When I was young, I \textbf{played} football. \trans{Jugaba fútbol}
    \item \textbf{Series of past events:} I \textbf{woke up}, \textbf{had} breakfast, and \textbf{went} to work.
\end{itemize}
\end{grammarbox}

\section{Regular Verbs: Add -ED}

Most verbs are \textbf{regular}. To form the past, add \textbf{-ed}.

\begin{grammarbox}[title=Regular Verbs Structure]
\textbf{Affirmative:} Subject + \textbf{verb + ed}

\textbf{Examples:}
\begin{itemize}
    \item I \textbf{worked} yesterday. \trans{Trabajé ayer}
    \item She \textbf{played} tennis. \trans{Ella jugó tenis}
    \item They \textbf{watched} TV. \trans{Miraron TV}
\end{itemize}
\end{grammarbox}

\subsection{Spelling Rules for -ED}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Rule} & \textbf{Example} & \textbf{Past Form} \\
\hline
Most verbs: add -ed & work, play, clean & worked, played, cleaned \\
\hline
Ends in -e: add -d only & live, like, move & lived, liked, moved \\
\hline
Ends in consonant + y: -ied & study, try, carry & studied, tried, carried \\
\hline
Ends in vowel + y: add -ed & play, enjoy, stay & played, enjoyed, stayed \\
\hline
One syllable CVC: double + ed & stop, plan, drop & stopped, planned, dropped \\
\hline
\end{tabular}
\caption{Spelling rules for regular past tense}
\end{table}

\subsection{Pronunciation of -ED}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Sound} & \textbf{After these sounds} & \textbf{Examples} \\
\hline
/t/ & voiceless: k, p, f, s, sh, ch & worked, stopped, laughed, washed \\
\hline
/d/ & voiced: b, g, l, m, n, r, v, z & played, lived, cleaned, opened \\
\hline
/id/ & after t or d & wanted, needed, decided, waited \\
\hline
\end{tabular}
\caption{Pronunciation of -ed endings}
\end{table}

\section{Irregular Verbs}

Many common verbs are \textbf{irregular}. They don't add -ed. You must learn them!

\begin{vocabbox}[title=Common Irregular Verbs]
\centering
\begin{tabular}{|l|l|l||l|l|l|}
\hline
\textbf{Infinitive} & \textbf{Past} & \textbf{Spanish} & \textbf{Infinitive} & \textbf{Past} & \textbf{Spanish} \\
\hline
be & was/were & ser/estar & have & had & tener \\
\hline
go & went & ir & do & did & hacer \\
\hline
come & came & venir & see & saw & ver \\
\hline
take & took & tomar & give & gave & dar \\
\hline
make & made & hacer & get & got & obtener \\
\hline
eat & ate & comer & drink & drank & beber \\
\hline
buy & bought & comprar & think & thought & pensar \\
\hline
write & wrote & escribir & read & read /red/ & leer \\
\hline
say & said & decir & tell & told & decir \\
\hline
know & knew & saber & meet & met & conocer \\
\hline
\end{tabular}
\end{vocabbox}

\section{Negative Sentences}

\begin{grammarbox}[title=Past Simple Negative]
\textbf{Structure:} Subject + \textbf{didn't} + \textbf{base verb}

\textbf{Important:} The main verb stays in the BASE FORM (no -ed)!

\textbf{Examples:}
\begin{itemize}
    \item I \textbf{didn't work} yesterday. \trans{No trabajé ayer}
    \item She \textbf{didn't go} to the party. \trans{Ella no fue a la fiesta}
    \item They \textbf{didn't see} the movie. \trans{No vieron la película}
\end{itemize}

\textbf{Common Error:}
\begin{itemize}
    \item \textcolor{red}{I didn't worked.} \textbf{X} (WRONG)
    \item \textcolor{green!60!black}{I didn't work.} \checkmark (CORRECT)
\end{itemize}
\end{grammarbox}

\section{Questions}

\begin{grammarbox}[title=Past Simple Questions]
\textbf{Structure:} \textbf{Did} + subject + \textbf{base verb}?

\textbf{Examples:}
\begin{itemize}
    \item \textbf{Did} you \textbf{work} yesterday? \trans{¿Trabajaste ayer?}
    \item \textbf{Did} she \textbf{go} to London? \trans{¿Fue ella a Londres?}
    \item What \textbf{did} you \textbf{do}? \trans{¿Qué hiciste?}
\end{itemize}

\textbf{Short Answers:}
\begin{itemize}
    \item Yes, I did. / No, I didn't.
    \item Yes, she did. / No, she didn't.
\end{itemize}

\textbf{Exception - Verb TO BE:}
\begin{itemize}
    \item \textbf{Was} he at home? (NOT: Did he be...?)
    \item \textbf{Were} they happy? (NOT: Did they be...?)
\end{itemize}
\end{grammarbox}

\section{Time Expressions}

\begin{vocabbox}[title=Common Past Time Expressions]
\begin{itemize}
    \item \textbf{yesterday} \trans{ayer}
    \item \textbf{last} night/week/month/year \trans{anoche/la semana pasada...}
    \item \textbf{ago:} 2 days ago, 3 weeks ago \trans{hace 2 días, hace 3 semanas}
    \item \textbf{in:} in 2020, in July \trans{en 2020, en julio}
    \item \textbf{when I was} young/a child \trans{cuando era joven/niño}
    \item \textbf{this morning} (if morning is finished) \trans{esta mañana}
\end{itemize}
\end{vocabbox}

\section{Practice Exercises}

\subsection{Exercise 1: Write the Past Form}
Write the past simple form of these verbs.

\begin{table}[h]
\centering
\begin{tabular}{|l|l||l|l|}
\hline
\textbf{Infinitive} & \textbf{Past} & \textbf{Infinitive} & \textbf{Past} \\
\hline
work & \underline{\hspace{3cm}} & go & \underline{\hspace{3cm}} \\
\hline
study & \underline{\hspace{3cm}} & see & \underline{\hspace{3cm}} \\
\hline
stop & \underline{\hspace{3cm}} & have & \underline{\hspace{3cm}} \\
\hline
play & \underline{\hspace{3cm}} & make & \underline{\hspace{3cm}} \\
\hline
live & \underline{\hspace{3cm}} & eat & \underline{\hspace{3cm}} \\
\hline
\end{tabular}
\end{table}

\subsection{Exercise 2: Complete the Sentences}
Fill in the blanks with the past simple form.

\begin{enumerate}
    \item I \underline{\hspace{3cm}} (visit) my grandparents last weekend.
    \item She \underline{\hspace{3cm}} (not/go) to work yesterday.
    \item They \underline{\hspace{3cm}} (watch) a movie on Saturday.
    \item \underline{\hspace{3cm}} you \underline{\hspace{3cm}} (have) a good time?
    \item He \underline{\hspace{3cm}} (be) very tired last night.
    \item We \underline{\hspace{3cm}} (not/see) Tom at the party.
    \item What time \underline{\hspace{3cm}} she \underline{\hspace{3cm}} (arrive)?
\end{enumerate}

\subsection{Exercise 3: Make Questions}
Transform these sentences into questions.

\begin{enumerate}
    \item You went to the cinema. $\rightarrow$ \underline{\hspace{6cm}}
    \item She bought a new car. $\rightarrow$ \underline{\hspace{6cm}}
    \item They lived in Paris. $\rightarrow$ \underline{\hspace{6cm}}
    \item He ate pizza for dinner. $\rightarrow$ \underline{\hspace{6cm}}
\end{enumerate}

\subsection{Exercise 4: Correct the Errors}
Find and correct the mistakes in these sentences.

\begin{enumerate}
    \item I didn't went to school yesterday.
    \item Did she came to the meeting?
    \item He didn't bought the book.
    \item They goed to London last year.
    \item Did you saw the movie?
\end{enumerate}

\subsection{Exercise 5: Complete the Story}
Fill in the blanks with the correct past simple form.

Last Saturday, I \underline{\hspace{2cm}} (wake) up early. I \underline{\hspace{2cm}} (have) breakfast and then I \underline{\hspace{2cm}} (go) to the gym. I \underline{\hspace{2cm}} (exercise) for an hour. After that, I \underline{\hspace{2cm}} (meet) my friends for lunch. We \underline{\hspace{2cm}} (eat) at a nice restaurant. In the afternoon, we \underline{\hspace{2cm}} (watch) a football match. My team \underline{\hspace{2cm}} (win)! In the evening, I \underline{\hspace{2cm}} (be) very tired, so I \underline{\hspace{2cm}} (stay) home and \underline{\hspace{2cm}} (read) a book.

\subsection{Exercise 6: Writing Task}
Write a paragraph (80-100 words) about what you did last weekend. Use at least 8 verbs in the past simple.

\begin{tcolorbox}[colback=white,height=7cm]
% Write about your last weekend...
\end{tcolorbox}

\section{Key Takeaways}
\begin{itemize}
    \item \textbf{Regular verbs} add -ed in the past (worked, played)
    \item \textbf{Irregular verbs} change form (go $\rightarrow$ went, see $\rightarrow$ saw)
    \item In \textbf{negatives and questions}, use \textbf{didn't/did} + base verb
    \item The main verb stays in BASE FORM after did/didn't
    \item Learn irregular verbs - they're very common!
    \item Use time expressions: yesterday, last week, ago
\end{itemize}
