
\subsection{what is your unique talent/ability or interest?}

I interest in mountain bike

I started to ride when I was 14 years old.

My first track in this sport was this one

Genearlly now I ride with my brother 

In this video we can see different lines

If you dont understand about the lines, is refer to the different paths that you can take in a track

https://www.youtube.com/watch?v=WHU7dyv9FT8




\subsection{computers and technology}

I am engineer in computer science, so I have a little knowledge about computers and technology.

I like coding and programming, now I am working a startup where I am integrate AI in cameras for different purposes.

To explain the architecture of any app's software, generally exist two principals components, the front-end and the back-end and database. The front-end is the part that the user can see and interact with, while the back-end is the part that works behind the scenes to make everything function properly. This two components work together to provide a seamless user experience.

Exict diferent frameworks and languages for the development of front-end and back-end, to explain better what is a framework, is a set of tools and libraries that help developers to create applications more easily and efficiently.

The most popular front-end framework now is React where is bassed in JavaScript. The own is Meta, the instagram. whatsapp and facebook have this frameworks.

The most popular back-end framework now is Django where is bassed in Python. The own is Django Software Foundation.

how is comunicateion between front-end and back-end?
The communication between the front-end and back-end typically occurs through APIs (Application Programming Interfaces). The front-end sends requests to the back-end server, which processes these requests, interacts with the database if necessary, and then sends back the appropriate responses to the front-end. This allows for dynamic content updates and user interactions without needing to reload the entire page. Common methods for this communication include RESTful APIs and GraphQL.






a) Could you .......... what you mean by that? 


b) Sorry, I’m not sure if I .........., what’s identity theft? 

c) Is everything .......... so far? 

d) So, if I .......... you correctly… 

e) Do you see what I ...........? 







