
\chapter{Vocabulary}

This chapter contains useful vocabulary words and phrases that were covered in previous classes. It includes definitions and example sentences to help you understand how to use them in context.

\section{Vocabulary}

\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Vocabulary} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
bursary & a scholarship to attend a college or university. & She received a bursary to help pay for her tuition. \\
\hline
career break & a period of time when someone stops working in their profession to rest or pursue other interests. & She took a career break to travel the world. \\
\hline
coach & a bus used for long-distance travel. & We took a coach to the countryside for the weekend. \\
\hline
commute & travel some distance between one's home and place of work on a regular basis. & I commute to the city every day for work. \\
\hline
crutches & a support used by a person who is unable to walk without assistance. & He had to use crutches after breaking his leg. \\
\hline
delightful & highly pleasing. & The cake she baked was absolutely delightful. \\
\hline
gap year & a year spent away from formal education, typically to travel or gain work experience. & He took a gap year before starting university. \\
\hline
get up to & to do something, often something bad. & What did you get up to at the party last night? \\
\hline
hangover & a painful physical condition following the excessive consumption of alcohol. & He had a terrible hangover after the party. \\
\hline
milestone & a significant event or stage in the development of something. & Graduating from college is a major milestone in life. \\
\hline
mishap & an unfortunate accident. & He had a mishap while hiking and sprained his ankle. \\
\hline
nicked & stolen (informal). & My car was nicked last night. \\
\hline
nightshifts & working during the night hours. & She works the nightshifts at the hospital. \\
\hline
outstanding & exceptionally good. & Her performance in the play was outstanding. \\
\hline
perk & a benefit or advantage. & One of the perks of the job is free lunch. \\
\hline
placement & the action of placing someone in a job or position. & The placement of the new employee was successful. \\
\hline
proofread & read (printer's proofs or other written or printed material) and mark any errors. & I need to proofread my essay before submitting it. \\
\hline
purse & a small bag used by women to carry money. & I am looking for my purse. \\
\hline
revision & the action of revising. & The document is under revision for accuracy. \\
\hline
sandal & a light shoe with either an openwork upper or straps attaching the sole to the foot. & She wore sandals to the beach. \\
\hline
scholarship & a grant or payment made to support a student's education, awarded on the basis of academic or other achievement. & She received a scholarship for her excellent grades. \\
\hline
shy & being reserved or having or showing nervousness or timidity in the company of other people. & She is shy and doesn't like speaking in public. \\
\hline
sightseeing & visiting famous places. & I went sightseeing in London. \\
\hline
stomach ache & dolor de estómago. & I have a stomach ache. \\
\hline
stuck & unable to move or be moved. & The car got stuck in the mud. \\
\hline
thin & having little, or too little, flesh or fat on the body. & The thin cat looked hungry. \\
\hline


\end{longtable}


\section{Informal Contractions}
\begin{longtable}{lll}
\textbf{Contraction} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
gonna & going to & I'm gonna study tonight \\
wanna & want to & Do you wanna go? \\
gotta & got to & I gotta leave now \\
kinda & kind of & It's kinda cold today \\
gimme & give me & Gimme a break \\
cya & see you later & Cya tomorrow \\
\end{longtable}
    
\section{British Slang \& Expressions}
\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Expression} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
a bloke & a man & There's a bloke at the door \\
a catch up & meet with someone to talk & We should have a catch up \\
a tenner & a 10 pound note & Can you lend me a tenner? \\
cheap as chips & very cheap & That shirt was cheap as chips \\
chop chop & hurry up & Come on, chop chop \\
dodgy & untrustworthy, suspicious & That looks a bit dodgy \\
fancy & I really want to & I fancy a coffee \\
feel right as rain & feel well again & I feel right as rain now \\
give someone a bell & call someone & Give me a bell later \\
grab some food & get something to eat & Let's grab some food \\
gutted & very disappointed & I was gutted to miss it \\
-ish & approximately & Meet me at 3-ish \\
knackered & very tired & I'm absolutely knackered \\
on and off & not consistently, not always & It's been raining on and off \\
peckish & snack & I'm feeling a bit peckish \\
shuteye & sleep & I need some shuteye \\
skint & broke, no money temporarily & I'm a bit skint this week \\
up for something & agree to do something & Are you up for a movie? \\
went off & exploded, made a loud noise & The alarm went off at 6am \\

\end{longtable}

\section{General Expressions}
\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Expression} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
bare-feet & without shoes & He walked barefoot on the grass \\
encourage & give support, confidence, or hope to (someone) & She encouraged me to apply \\
fire away & is used to give someone permission to ask questions or make comments & Fire away, I'm ready \\
get back in touch & contact someone again & I'll get back in touch next week \\
hop & jump on one foot & The kids like to hop around \\
pay attention & focus on something & Please pay attention to the instructions \\
ran me over & hit me with a vehicle & A car ran me over \\
ran out of & have no more of (something) & I've run out of time \\
stay out & remain outside & We decided to stay out late \\
take it easy & relax & Take it easy this weekend \\
wander around & walk around without a specific purpose & We wandered around the city \\
\end{longtable}


\section{Order Words/Phrases}
\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Word/Phrase} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
after that & following that; subsequently & After that, we can make a decision \\
\hline
finally & at the end; lastly & Finally, we will summarize our findings \\
\hline
firstly & used to introduce the first point or reason in a list & Firstly, we need to gather information \\
\hline
following that & after that; subsequently & Following that, we will review the plan \\
\hline
next & after that; then & Next, we will discuss the results \\
\hline
secondly & used to introduce the second point or reason in a list & Secondly, we need to analyze the data \\
\hline


\end{longtable} 

\section{Collocations}
\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Preposition} & \textbf{Usage} & \textbf{Example} \\
\hline
\endhead
at & used with "the weekend" & At the weekend \\
\hline
in & used with parts of the day & In the morning/afternoon/evening \\
\hline
on & used with days & On Saturday \\
\hline
\end{longtable} 








