\chapter{Past Habits: Would and Used To}

\begin{center}
\begin{tabular}{ccc}
\cefrlevel{B1-B2} & \textbf{Study Time:} 2-3 hours & \textbf{Difficulty:} ⭐⭐⭐☆☆
\end{tabular}
\end{center}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item How to use ``used to'' for past habits and states
    \item How to use ``would'' for past repeated actions
    \item The difference between ``used to'' and ``would''
    \item Stative verbs that cannot be used with ``would''
\end{itemize}

\section{Reading Context}
\begin{readingbox}[title=Dialogue: Childhood Memories]
\textbf{Grandpa:} This photo brings back memories. I \textbf{used to} live in a small village.\\
\textbf{Grandson:} Really? What was it like?\\
\textbf{Grandpa:} It was quiet. We didn't have TV. Every evening, my father \textbf{would} tell us stories.\\
\textbf{Grandson:} That sounds nice. Did you play video games?\\
\textbf{Grandpa:} No! We \textbf{used to} play outside all day. In the summer, we \textbf{would} go swimming in the river.\\
\textbf{Grandson:} I \textbf{used to} be afraid of swimming.\\
\textbf{Grandpa:} Me too, at first. But I loved it later.
\end{readingbox}

\begin{comparisonbox}[title=🇪🇸 "Solía" en Español vs USED TO/WOULD en Inglés]
\textbf{⚠️ Español tiene "solía" + infinitivo para todo, inglés distingue estados vs acciones}

\textbf{1. Español - "solía" universal:}
\begin{itemize}
    \item "Solía \textit{vivir} en Madrid" (estado)
    \item "Solía \textit{jugar} fútbol" (acción repetida)
    \item "Solía \textit{ser} tímido" (estado)
    \item Todos usan la misma estructura
\end{itemize}

\textbf{2. Inglés - USED TO para estados, USED TO/WOULD para acciones:}
\begin{itemize}
    \item Estados (be, have, live, know) → SOLO "used to"
    \begin{itemize}
        \item I \textbf{used to live} in Madrid ✓
        \item I \textbf{would live} in Madrid ❌
    \end{itemize}
    \item Acciones repetidas (play, go, eat) → "used to" O "would"
    \begin{itemize}
        \item I \textbf{used to play} football ✓
        \item I \textbf{would play} football ✓ (ambas correctas)
    \end{itemize}
\end{itemize}

\textbf{3. Imperfecto español → USED TO/WOULD inglés:}
\begin{itemize}
    \item "Cuando era niño, \textit{jugaba} fútbol" = When I was young, I \textbf{used to play} / \textbf{would play}
    \item "Cuando era niño, \textit{vivía} en Madrid" = When I was young, I \textbf{used to live} (NO "would live")
\end{itemize}

\textbf{💡 Truco: ¿Puedo decir "estaba + gerundio"?}
\begin{itemize}
    \item Si NO → Es estado → SOLO "used to"
    \item Si SÍ → Es acción → "used to" O "would"
\end{itemize}

\textbf{Ejemplos:}
\begin{itemize}
    \item "Vivir" → NO puedo decir "estaba viviendo" habitualmente → used to ✓, would ❌
    \item "Jugar" → SÍ puedo decir "estaba jugando" → used to ✓, would ✓
\end{itemize}
\end{comparisonbox}

\section{Key Concepts: Past Habits}

We use two structures to talk about things we did in the past but don't do now.

\begin{vocabbox}[title=Used To vs. Would]
\begin{itemize}
    \item \textbf{Used to + Verb:} For past habits AND states.
    \item \textbf{Would + Verb:} ONLY for past repeated actions.
\end{itemize}
\end{vocabbox}

\section{Grammar Focus: The Rules}

\begin{grammarbox}[title=Rule 1: Action Verbs]
For repeated actions (play, go, visit, eat), you can use both:
\begin{itemize}
    \item I \textbf{used to} play football. \checkmark
    \item I \textbf{would} play football. \checkmark
\end{itemize}
\end{grammarbox}

\begin{grammarbox}[title=Rule 2: Stative Verbs]
For states (be, have, live, love, know), use ONLY \textbf{used to}:
\begin{itemize}
    \item I \textbf{used to} be shy. \checkmark
    \item I \textbf{would} be shy. \textbf{X} (Incorrect)
    \item I \textbf{used to} have a car. \checkmark
    \item I \textbf{would} have a car. \textbf{X} (Incorrect)
\end{itemize}
\end{grammarbox}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Stative Verbs (Use 'Used to')} & \textbf{Action Verbs (Use 'Used to' or 'Would')} \\
\hline
Be, Have, Live, Love, Like, Know & Go, Play, Visit, Walk, Eat, Read \\
\hline
\end{tabular}
\caption{Stative vs. Action Verbs}
\end{table}

\section{Practice Exercises}

\subsection{Exercise 1: Used To or Would?}
Mark if the sentence is Correct (C) or Incorrect (I).

\begin{enumerate}
    \item I would live in Paris. (\_\_)
    \item She used to have long hair. (\_\_)
    \item Every Sunday, we would visit grandma. (\_\_)
    \item He would know the answer. (\_\_)
\end{enumerate}

\subsection{Exercise 2: Fill in the Blanks}
Use \textit{would} if possible. If not, use \textit{used to}.

\begin{enumerate}
    \item I \underline{\hspace{3cm}} (be) a teacher.
    \item Every summer, we \underline{\hspace{3cm}} (go) to the beach.
    \item She \underline{\hspace{3cm}} (love) chocolate.
    \item He \underline{\hspace{3cm}} (play) the guitar every night.
\end{enumerate}

\subsection{Exercise 3: Writing Task}
Write a short paragraph (60-80 words) about your childhood. Use \textbf{used to} for states and \textbf{would} for repeated actions.

\begin{tcolorbox}[colback=white,height=5cm]
% Write here...
\end{tcolorbox}

\section{Key Takeaways}
\begin{itemize}
    \item \textbf{Used to} works for everything (habits and states).
    \item \textbf{Would} only works for actions (running, playing), not states (being, having).
    \item Negative: \textbf{didn't use to} (no 'd').
    \item Question: \textbf{Did you use to...?}
\end{itemize}

\section{Online Practice}
\begin{tcolorbox}[colback=blue!5,colframe=blue!40!black,title=Resources to Practice]
Online PracticeReinforce what you have learned with these interactive exercises:
\begin{itemize}
    \item \textbf{Used To / Would:} \url{https://www.perfect-english-grammar.com/used-to.html}
    \item \textbf{Past Habits Exercises:} \url{https://test-english.com/grammar-points/b1/used-to-would/}
    \item \textbf{Interactive Practice:} \url{https://learnenglish.britishcouncil.org/grammar/intermediate-to-upper-intermediate/used-to}
    \item \textbf{Past Habits Quiz:} \url{https://www.englishpage.com/verbpage/usedto.html}
\end{itemize}
\end{tcolorbox}





