\chapter{Accommodation and Housing Problems}

\begin{center}
\begin{tabular}{ccc}
\cefrlevel{B1-B2} & \textbf{Study Time:} 2-3 hours & \textbf{Difficulty:} ⭐⭐⭐☆☆
\end{tabular}
\end{center}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item Vocabulary for different types of accommodation
    \item How to describe your living space and furniture
    \item Common housing problems (leaks, stains, scratches, etc.)
    \item How to make complaints and request repairs
    \item Describing places and landscapes effectively
\end{itemize}

\section{Types of Accommodation}

When living abroad or travelling, you might encounter different types of accommodation.

\begin{vocabbox}[title=Accommodation Types]
\begin{itemize}
    \item \keyterm{Hostel} - Budget accommodation with shared rooms (dormitories). Can be noisy ("snoring") and lack privacy ("no intimacy"). Often has bad smells due to many people sharing spaces. Good for meeting people and saving money.
    \item \keyterm{Flat-share} - Sharing an apartment with other people. You usually have your own room but share the kitchen and bathroom. Good for independence while splitting costs.
    \item \keyterm{Homestay} - Living with a local family. Good for practicing language and experiencing local culture, but rules might be strict. May have issues like noisy children.
    \item \keyterm{Hall of Residence} (Student Hall) - Accommodation provided by universities. Often small rooms but social atmosphere. Good for students.
\end{itemize}
\end{vocabbox}

\subsection{Accommodation Pros and Cons}

\begin{table}[h]
\centering
\begin{tabular}{|l|p{4cm}|p{4cm}|}
\hline
\textbf{Type} & \textbf{Pros} & \textbf{Cons} \\
\hline
Hostel & Cheap, social, central locations & Noisy (snoring), no privacy (intimacy), sometimes bad smell \\
\hline
Flat-share & Independent, shared costs, own room & Depends on flatmates, shared facilities \\
\hline
Homestay & Language practice, cultural experience, meals included & Strict rules, less privacy, family dynamics (noisy children) \\
\hline
Hall of Residence & University community, close to campus, social & Small rooms, shared facilities, can be noisy \\
\hline
\end{tabular}
\end{table}

\section{Describing Your Accommodation}

\begin{readingbox}[title=Sample Description: My Student Room]
I live in a \textbf{student castle}, which is located in Oxford. My room in this accommodation is small but \textbf{cozy}. It has a bed, a desk, and a little toilet.
For storage, I have a \textbf{wardrobe} where I hang my clothes and store my shoes. There is also a \textbf{shelf} near my desk, which I use to keep folders and other things. On the wall, there is a \textbf{board} to pin notes and photos.
Sometimes I listen to music from my neighbor, who practices the saxophone; it is actually quite relaxing to hear.
\end{readingbox}

\subsection{Useful Vocabulary}
\begin{itemize}
    \item \textbf{Cozy} - Small, warm, and comfortable.
    \item \textbf{Wardrobe} - A large cupboard for hanging clothes.
    \item \textbf{Shelf} - A flat board to put things on.
    \item \textbf{Pinboard/Bulletin board} - A board for displaying notices.
\end{itemize}

\section{Common Housing Problems}

\subsection{Defects and Damage}

\begin{table}[h]
\centering
\begin{tabular}{|l|p{5cm}|p{5cm}|}
\hline
\textbf{Problem} & \textbf{Definition} & \textbf{Collocations \& Examples} \\
\hline
\textbf{Leak} & Liquid escaping through a hole or crack. & \textit{Abundant leak, fix a leak.}\\ "The pipe has a leak." \\
\hline
\textbf{Scratch} & A slight cut or mark on a surface. & \textit{Light/Deep/Fine/Surface/Visible scratch.}\\ "The cat left scratches on the door." \\
\hline
\textbf{Dent} & A hollow mark in a hard surface caused by a blow. & \textit{Small/Deep/Minor/Large dent.}\\ "The hailstorm caused a dent on the roof." \\
\hline
\textbf{Stain} & A colored patch that is difficult to remove. & \textit{Blood/Ink/Oil/Wine stain.}\\ "Red wine caused a stain on the carpet." \\
\hline
\textbf{Rip/Tear} & A hole caused by pulling apart (fabric/paper). & \textit{Jagged rip, torn muscle.}\\ "There is a rip in the sofa." \\
\hline
\textbf{Crack} & A line on the surface where it has split without breaking apart. & "There is a crack in the wall." \\
\hline
\textbf{Chip} & A small piece broken off the edge or surface. & \textit{Chipped paint, chipped plate.}\\ "The cup is chipped." \\
\hline
\end{tabular}
\end{table}

\subsection{Extended Damage Vocabulary}

\begin{table}[h]
\centering
\begin{tabular}{|l|p{10cm}|}
\hline
\textbf{Scratch Collocations} & \\
\hline
\textit{leave a scratch} & The knife left a scratch on the plate. \\
\textit{cause a scratch} & The cat caused scratches on the furniture. \\
\textit{repair a scratch} & I had to repair the scratch on my car. \\
\textit{remove a scratch} & I tried to remove the scratch from the table. \\
\textit{cover a scratch} & She used a sticker to cover the scratch on her laptop. \\
\hline
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{|l|p{10cm}|}
\hline
\textbf{Dent Collocations} & \\
\hline
\textit{make a dent} & The high pavement made a dent in my car. \\
\textit{leave a dent} & The car left a dent on the garage door. \\
\textit{cause a dent} & The hailstorm caused a dent on my roof. \\
\textit{repair a dent} & I need to repair the dent on my car door. \\
\textit{fix a dent} & The mechanic will fix the dent in my car. \\
\hline
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{|l|p{10cm}|}
\hline
\textbf{Stain Collocations} & \\
\hline
\textit{blood stain} & The shirt has a blood stain on it. \\
\textit{ink stain} & There's an ink stain on the carpet. \\
\textit{oil stain} & The garage floor has several oil stains. \\
\textit{wine stain} & Red wine stains are difficult to remove. \\
\textit{wash out/off a stain} & I need to wash off this stain. \\
\textit{cause a stain} & The red wine caused a stain on the carpet. \\
\textit{remove a stain} & I used a special cleaner to remove the stain from my shirt. \\
\hline
\end{tabular}
\end{table}

\subsection{Additional Damage Vocabulary}

\begin{vocabbox}[title=More Damage and Problems]
\begin{itemize}
    \item \textbf{Leaking} - When water or other liquid escapes from a container or pipe
    \item \textbf{Smashed} - Broken into many pieces
    \item \textbf{Cracked} - Having a line on the surface where it has broken but not completely separated
    \item \textbf{Run out} - When something is completely used up and there is no more left
    \item \textbf{Dead} (batteries) - No longer functioning
    \item \textbf{Chipped} - Having small pieces broken off the edge or surface
    \item \textbf{Power cut} - A temporary interruption in the supply of electricity
    \item \textbf{Come off} - To become detached or separated from something
    \item \textbf{Fall off} - To drop down from a higher position
    \item \textbf{Break down} - To stop functioning properly (machines, vehicles)
    \item \textbf{Get crushed} - To be damaged by being pressed or squeezed with force
    \item \textbf{Crash} - Colliding with something violently
\end{itemize}
\end{vocabbox}

\subsection{Verbs for Damage and Repair}

\begin{itemize}
    \item \textbf{To cause/leave} a scratch/dent/stain: "The knife \textbf{left} a scratch on the plate."
    \item \textbf{To fix/repair/mend} a problem: "I need to \textbf{repair} the dent."
    \item \textbf{To remove} a stain/scratch: "I used a cleaner to \textbf{remove} the stain."
    \item \textbf{To cover} up: "She used a sticker to \textbf{cover} the scratch."
\end{itemize}

\section{Making Complaints and Requests}

When you have a problem, you need to use polite but firm language to ask for a solution.

\begin{readingbox}[title=Dialogue: Reporting a Problem]
\textbf{Tenant:} Hello, I have a request. I am staying in a homestay, but I have a problem with the family.\\
\textbf{Agent:} I'm sorry to hear that. What is the problem?\\
\textbf{Tenant:} The children are very noisy, and I can't work in my room. I tried to talk to them, but they didn't listen.\\
\textbf{Agent:} I understand. We want to make sure you have a comfortable stay. Let me check if we have another homestay available.\\
\textbf{Tenant:} Thank you, I would appreciate that. If possible, I would like to move as soon as possible, or I would like a \textbf{reduction in my rent}.
\end{readingbox}

\subsection{Complaint Vocabulary}

\begin{vocabbox}[title=Key Words for Complaints]
\begin{itemize}
    \item \keyterm{Attention} - Focus or care given to something
    \item \keyterm{Wrong} - Not correct or not working properly
    \item \keyterm{Grateful} - Feeling or showing appreciation
    \item \keyterm{Disappointed} - Sad because something didn't meet expectations
    \item \keyterm{Dissatisfaction} - Feeling of not being satisfied
    \item \keyterm{Inconvenience} - A problem or difficulty that causes trouble
    \item \keyterm{Sort (out)} - To fix or resolve a problem
    \item \keyterm{Fuss} - Unnecessary excitement or worry about something
    \item \keyterm{Legal} - Related to the law
    \item \keyterm{Incompetence} - Lack of ability to do something properly
    \item \keyterm{Misunderstanding} - A failure to understand something correctly
    \item \keyterm{Complain} - To express dissatisfaction or annoyance
    \item \keyterm{Angry} - Feeling or showing strong displeasure
\end{itemize}
\end{vocabbox}

\subsection{Example Complaint Scenario}

\begin{tcolorbox}[colback=red!5, colframe=red!30, title=Sample Complaint Letter Structure]
\textbf{Problem:} Light not working in hallway and kitchen

\textbf{Key Phrases:}
\begin{enumerate}
    \item \textbf{State the problem:} "The light in the hallway and in the kitchen is not working."
    \item \textbf{Request action:} "Please fix this as soon as possible."
    \item \textbf{Suggest solution:} "Could you send an electrician to fix it?"
    \item \textbf{Express concern:} "This is taking a long time to fix."
    \item \textbf{Request compensation:} "I would like a reduction in the rent due to this inconvenience."
\end{enumerate}
\end{tcolorbox}

\subsection{Useful Phrases for Complaints}
\begin{itemize}
    \item "I have a request regarding..."
    \item "I would appreciate it if you could..."
    \item "I would like to change/move..."
    \item "There seems to be a problem with..."
    \item "Could you please send an electrician/plumber?"
    \item "I assure you this is a serious concern."
    \item "Would you mind sorting this out?"
    \item "If this continues, I may need to consider legal action."
    \item "I'm grateful for your attention to this matter."
    \item "I'm disappointed with the incompetence shown."
    \item "There seems to be a misunderstanding."
\end{itemize}

\subsection{Exercise: Role Play - Making a Complaint}

\textbf{Scenario:} You are staying in accommodation and have a problem. Practice this dialogue with a partner:

\begin{tcolorbox}[colback=yellow!5, colframe=yellow!30, title=Sample Dialogue]
\textbf{A:} Hello\\
\textbf{B:} Hi, may I help you?\\
\textbf{A:} Yes, I have a request. I am staying in a homestay, but I have a problem with my host family.\\
\textbf{B:} Oh, I'm sorry to hear that. What is the problem?\\
\textbf{A:} My host family's children are very noisy, and I can't work in my room. I tried to talk with them, but they didn't listen to me.\\
\textbf{B:} I understand. We want to make sure you have a comfortable stay. Let me check if we have another homestay available for you.\\
\textbf{A:} Thank you, I would appreciate that. And if it's possible, I would like to change my homestay as soon as possible, or I want a reduction in my rent.\\
\textbf{B:} I will do my best to accommodate your request. Please give me a moment to check our records.\\
\textbf{A:} Thank you very much for your help.\\
\textbf{B:} You're welcome. I will get back to you shortly with an update.
\end{tcolorbox}

\subsection{Common Repairs}
\begin{enumerate}
    \item Fix a leak
    \item Paint the walls
    \item Change the lighting
    \item Replace broken tiles
    \item Remodel the kitchen
    \item Landscape the garden
    \item Restore old furniture
    \item Lay wooden floorboards
\end{enumerate}

\section{The Causative: "Have Something Done"}

When you arrange for someone else to do something for you (not doing it yourself), use the \textbf{causative structure}:

\begin{grammarbox}[title=Have Something Done Structure]
\textbf{Structure:} to have + \textbf{object} + \textbf{past participle}

\textbf{Formula:} Subject + have/has/had + object + V3 (past participle)

\textbf{Examples:}
\begin{itemize}
    \item I \textbf{have} my car \textbf{repaired}. \trans{Hago que reparen mi coche.}
    \item She \textbf{has} her hair \textbf{cut} every month. \trans{Ella se corta el pelo cada mes.}
    \item We \textbf{had} the house \textbf{painted} last year. \trans{Pintamos la casa el año pasado (alguien más lo hizo).}
\end{itemize}

\textbf{Key Point:} This structure shows that \textit{someone else} does the action for you, not you doing it yourself.
\end{grammarbox}

\subsection{Tenses with "Have Something Done"}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Tense} & \textbf{Structure} & \textbf{Example} \\
\hline
Present Simple & have/has + object + V3 & I have my car serviced annually. \\
\hline
Present Continuous & am/is/are having + object + V3 & We are having the walls painted now. \\
\hline
Past Simple & had + object + V3 & I had my keys made yesterday. \\
\hline
Future Simple & will have + object + V3 & I will have my hair cut tomorrow. \\
\hline
Present Perfect & have/has had + object + V3 & I have had my bike repaired twice. \\
\hline
Modal verbs & modal + have + object + V3 & You should have it checked. \\
\hline
\end{tabular}
\end{table}

\subsection{Practice: Have Something Done}

\textbf{Exercise 1:} Rewrite the sentences using "have something done":

\begin{enumerate}
    \item I won't fix the leak in the kitchen myself. \\
    \textbf{Answer:} I will have the leak fixed.
    
    \item They aren't going to paint their house themselves. \\
    \textbf{Answer:} They are going to have it painted.
    
    \item He shouldn't replace the broken tile himself as he will do more damage. \\
    \textbf{Answer:} He should have it replaced.
    
    \item We didn't have time to landscape the garden ourselves. \\
    \textbf{Answer:} We had it landscaped.
\end{enumerate}

\subsection{More Examples}

\begin{tcolorbox}[colback=blue!5, colframe=blue!30, title=Have Something Done Examples]
\begin{itemize}
    \item We \textbf{are having} the bedroom walls \textbf{painted} at the moment.
    \item My house is finally cool. I \textbf{had} air conditioning \textbf{installed} yesterday.
    \item I lost my keys. I will \textbf{have} to get a new set \textbf{made}.
    \item Your garden looks like a jungle. When was the last time you \textbf{had} the grass \textbf{cut}?
    \item \textbf{Have} you \textbf{had} the bathroom \textbf{remodeled}? It looks different.
    \item How often do you \textbf{have} the walls \textbf{painted}?
\end{itemize}
\end{tcolorbox}

\subsection{Exercise 2: Describing Accommodation Problems}

Look at the description below and identify all the problems. Then say how you would fix them using "have something done":

\begin{readingbox}[title=Problem Accommodation]
\textbf{Current condition:}
\begin{itemize}
    \item The sofa is torn and damaged
    \item The floor is dirty and stained
    \item The carpet is dirty and poorly maintained
    \item The walls have several deep cracks
    \item There is a water leak
    \item The table is dirty
    \item The pipes on the ceiling are exposed
    \item The pillows are worn out and need to be replaced
\end{itemize}
\end{readingbox}

\textbf{Solutions using "have something done":}
\begin{itemize}
    \item I will \textbf{have} the sofa \textbf{restored}.
    \item I will \textbf{have} the floor \textbf{cleaned}.
    \item I will \textbf{have} the carpet \textbf{professionally cleaned}.
    \item I will \textbf{have} the wall cracks \textbf{repaired}.
    \item I will \textbf{have} the water leak \textbf{fixed}.
    \item I will \textbf{have} the table \textbf{cleaned}.
    \item I will \textbf{have} the ceiling pipes \textbf{covered}.
    \item I will \textbf{have} the pillows \textbf{replaced}.
    \item I will \textbf{have} plants \textbf{added} to the room.
    \item I will \textbf{have} light fixtures \textbf{installed}.
\end{itemize}

\subsection{Writing Task: Accommodation Improvement Plan}

Write two paragraphs:
\begin{enumerate}
    \item \textbf{Paragraph 1:} Describe the current condition of your accommodation (or an imaginary one) with problems.
    \item \textbf{Paragraph 2:} Propose solutions using "have something done" structure.
\end{enumerate}

\begin{tcolorbox}[colback=green!5, colframe=green!30, title=Sample Essay]
\textbf{Current Condition:}

In this accommodation, there are several issues that need to be addressed. The sofa is torn and damaged, making it uncomfortable to sit on. The floor is dirty and stained, which detracts from the overall cleanliness of the room. Additionally, the carpet is poorly maintained and requires thorough cleaning. The walls have several deep cracks that need to be repaired to prevent further damage. There is also a water leak that needs immediate attention to avoid potential flooding. The table is dirty and needs to be cleaned, while the exposed pipes on the ceiling give the room an unfinished look. Lastly, the pillows are worn out and should be replaced for better comfort.

\textbf{Proposed Solutions:}

To improve the living conditions in this accommodation, I plan to have several repairs and maintenance tasks completed. I will have the sofa restored to fix the tears and damage, making it comfortable again. The floor will be cleaned to remove stains and improve its appearance. I will also have the carpet professionally cleaned to ensure it is hygienic and well-maintained. The deep cracks in the walls will be repaired to prevent further deterioration. The water leak will be fixed promptly to avoid any potential flooding issues. I will have the table cleaned to enhance its usability, and the exposed pipes on the ceiling will be covered to give the room a more polished look. Finally, I will have new pillows replaced to ensure comfort during my stay. Additionally, I will have plants added to the room, windows installed in the left wall, and a new wardrobe added to improve storage space.
\end{tcolorbox}

\section{Describing Places and Landscapes}

When describing a beautiful place, use strong adjectives to create a picture.

\begin{readingbox}[title=Sample Description: Torres del Paine]
One of the most beautiful places I have ever visited is \textbf{Torres del Paine}. It features rocky, \textbf{snow-covered mountains} located in the south of Chile.
The view of the mountains is \textbf{spellbinding}. It is truly \textbf{breathtaking} to see the peaks against the clear blue sky.
Other features include:
\begin{itemize}
    \item Freshwater lakes
    \item Grassy meadows and rolling hills
    \item Mysterious dense forests
    \item Picturesque churches (e.g., on an island like in Bled)
\end{itemize}
\end{readingbox}

\subsection{Adjectives for Places}
\begin{itemize}
    \item \textbf{Breathtaking} - Extremely beautiful/impressive.
    \item \textbf{Spellbinding} - Holding your attention completely; fascinating.
    \item \textbf{Picturesque} - Visually attractive (like a picture).
    \item \textbf{Dense} (forest) - Thick, difficult to see through.
    \item \textbf{Rolling} (hills) - Gently rising and falling.
\end{itemize}
