\chapter{Improvement, Action, and Verb Patterns}

\begin{center}
\begin{tabular}{ccc}
\cefrlevel{B1-B2} & \textbf{Study Time:} 3-4 hours & \textbf{Difficulty:} ⭐⭐⭐⭐☆
\end{tabular}
\end{center}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item Vocabulary related to making improvements and taking action
    \item Collocations with "give"
    \item Grammatical patterns of the verb "give"
    \item How to describe change and progress
\end{itemize}

\section{Reading Context}
\begin{readingbox}[title=Dialogue: Improving Team Performance]
\textbf{Manager:} Our sales numbers are down. We need to \textbf{boost} our performance.\\
\textbf{Team Lead:} I agree. Maybe we should \textbf{give} the new marketing strategy \textbf{a try}.\\
\textbf{Manager:} Good idea. It might \textbf{strengthen} our brand presence.\\
\textbf{Team Lead:} Also, the team seems tired. They need to \textbf{recharge their batteries}.\\
\textbf{Manager:} True. Let's \textbf{give them a break} this weekend.\\
\textbf{Team Lead:} Thanks. I'll \textbf{give you a hand} with the new schedule.\\
\textbf{Manager:} Great. Let's \textbf{give it our best} shot next month.
\end{readingbox}

\section{Key Concepts: Change and Improvement}

\begin{vocabbox}[title=Improvement Vocabulary]
\begin{itemize}
    \item \keyterm{Boost} (verb): To help or encourage something to increase or improve.
    \begin{itemize}
        \item \trans{Impulsar, mejorar}
    \end{itemize}
    \item \keyterm{Strengthen} (verb): To make something stronger or more effective.
    \begin{itemize}
        \item \trans{Fortalecer}
    \end{itemize}
    \item \keyterm{Enhance} (verb): To improve the quality, value, or extent of something.
    \begin{itemize}
        \item \trans{Mejorar, realzar}
    \end{itemize}
    \item \keyterm{Recharge batteries} (idiom): To rest and recover energy levels.
    \begin{itemize}
        \item \trans{Recargar las pilas}
    \end{itemize}
\end{itemize}
\end{vocabbox}

\section{Grammar Focus: Verb Patterns with "Give"}

The verb \textit{give} is used in many fixed expressions and patterns in English.

\begin{grammarbox}[title=Pattern 1: Give + It + a + Noun]
Used to express attempting or trying something.

\textbf{Expressions:}
\begin{itemize}
    \item \textbf{Give it a try} \trans{Intentarlo}
    \item \textbf{Give it a go} \trans{Intentarlo (British)}
    \item \textbf{Give it a shot} \trans{Darle una oportunidad (American)}
    \item \textbf{Give it some thought} \trans{Pensarlo}
\end{itemize}
\end{grammarbox}

\begin{grammarbox}[title=Pattern 2: Give + Someone + Something]
Used when giving something to a person (often abstract).

\textbf{Expressions:}
\begin{itemize}
    \item \textbf{Give someone a hand} \trans{Echar una mano / Ayudar}
    \item \textbf{Give someone a lift} \trans{Llevar a alguien (en coche)}
    \item \textbf{Give someone a call} \trans{Llamar a alguien}
    \item \textbf{Give someone a break} \trans{Dar un respiro}
\end{itemize}
\end{grammarbox}

\section{Collocations Reference}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Verb} & \textbf{Common Collocations} \\
\hline
Boost & confidence, morale, sales, performance, economy \\
\hline
Strengthen & relationship, bond, muscles, position, argument \\
\hline
Improve & skills, health, quality, situation, performance \\
\hline
Enhance & experience, quality, appearance, flavor, value \\
\hline
Weaken & position, argument, immune system, currency \\
\hline
\end{tabular}
\caption{Common collocations with improvement verbs}
\end{table}

\section{Practice Exercises}

\subsection{Exercise 1: Complete with "Give" Expressions}

\begin{enumerate}
    \item I'm not sure if I'll like yoga, but I'll \underline{\hspace{3cm}} (try it).
    \item Can you \underline{\hspace{3cm}} (help me) with my homework?
    \item I need to \underline{\hspace{3cm}} (think about it) before I make a decision.
    \item Do you want me to \underline{\hspace{3cm}} (drive you) to the station?
\end{enumerate}

\subsection{Exercise 2: Vocabulary in Context}
Fill in the blanks with: \textit{boost, strengthen, recharge, practical}.

\begin{enumerate}
    \item We need to \underline{\hspace{2cm}} our marketing to get more customers.
    \item This course offers \underline{\hspace{2cm}} advice for daily life.
    \item I'm going on vacation to \underline{\hspace{2cm}} my batteries.
    \item Exercise helps to \underline{\hspace{2cm}} your muscles.
\end{enumerate}

\subsection{Exercise 3: Discuss Improvements}
Talk about something you want to improve in your life. Use at least 3 words from the vocabulary list.

\begin{tcolorbox}[colback=white,height=4cm]
% Write your answer here
\end{tcolorbox}

\section{Key Takeaways}
\begin{itemize}
    \item Use "boost" and "strengthen" to talk about positive changes.
    \item "Give it a try/go/shot" means to attempt something.
    \item "Give someone a hand" means to help someone.
    \item Learn collocations together with verbs for natural speech.
\end{itemize}

\section{Online Practice}
\begin{tcolorbox}[colback=blue!5,colframe=blue!40!black,title=Resources to Practice]
Online PracticeReinforce what you have learned with these interactive exercises:
\begin{itemize}
    \item \textbf{Collocations with GIVE:} \url{https://www.englishclub.com/ref/Collocations/give/}
    \item \textbf{Verb Patterns Practice:} \url{https://test-english.com/grammar-points/b1-b2/verb-patterns-gerund-infinitive/}
    \item \textbf{Change Vocabulary:} \url{https://www.englishclub.com/vocabulary/}
    \item \textbf{Business Collocations:} \url{https://www.businessenglishpod.com/category/collocations/}
\end{itemize}
\end{tcolorbox}
