\chapter{Verb Patterns: Gerunds and Infinitives}

\begin{center}
\begin{tabular}{ccc}
\cefrlevel{B2-C1} & \textbf{Study Time:} 3-4 hours & \textbf{Difficulty:} ⭐⭐⭐⭐☆
\end{tabular}
\end{center}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item When to use the Gerund (verb + -ing)
    \item When to use the Infinitive (to + verb)
    \item Common verbs that follow each pattern
    \item How to give recommendations and talk about plans
\end{itemize}

\section{Reading Context}
\begin{readingbox}[title=Dialogue: Weekend Plans]
\textbf{Anna:} What do you \textbf{plan to do} this weekend?\
\textbf{Ben:} I \textbf{want to go} to the cinema. I \textbf{suggest watching} the new Marvel movie.\
\textbf{Anna:} I \textbf{avoid going} to the cinema on Saturdays. It's too busy.\
\textbf{Ben:} Really? I \textbf{enjoy seeing} movies with a big crowd.\
\textbf{Anna:} I \textbf{prefer to stay} home. I \textbf{decided to cook} a nice dinner.\
\textbf{Ben:} That sounds good. Do you \textbf{need to buy} ingredients?\
\textbf{Anna:} Yes. I \textbf{hope to find} fresh fish at the market.
\end{readingbox}

\begin{comparisonbox}[title=🇪🇸 Gerundios e Infinitivos: Pesadilla para Hispanohablantes]
\textbf{⚠️ En español usamos infinitivo casi siempre, en inglés depende del verbo anterior}

\textbf{1. Español - mayormente infinitivo:}
\begin{itemize}
    \item "Me gusta \textit{bailar}" (infinitivo)
    \item "Quiero \textit{comer}" (infinitivo)
    \item "Evito \textit{hablar}" (infinitivo)
    \item "Termino de \textit{trabajar}" (infinitivo)
\end{itemize}

\textbf{2. Inglés - algunos verbos + gerundio (-ing):}
\begin{itemize}
    \item "I enjoy \textbf{dancing}" (NO "to dance")
    \item "I finish \textbf{working}" (NO "to work")
    \item "I avoid \textbf{speaking}" (NO "to speak")
\end{itemize}

\textbf{3. Inglés - otros verbos + infinitivo (to + verb):}
\begin{itemize}
    \item "I want \textbf{to eat}" (NO "eating")
    \item "I hope \textbf{to see}" (NO "seeing")
    \item "I decide \textbf{to go}" (NO "going")
\end{itemize}

\textbf{4. NO hay regla lógica - hay que memorizar:}
\begin{center}
\textbf{LIKE} puede usar AMBOS: "I like \textbf{dancing}" = "I like \textbf{to dance}"\\
\textbf{ENJOY} solo gerundio: "I enjoy \textbf{dancing}" ✓ / "I enjoy \textbf{to dance}" ❌\\
\textbf{WANT} solo infinitivo: "I want \textbf{to dance}" ✓ / "I want \textbf{dancing}" ❌
\end{center}

\textbf{💡 Truco: Verbos de "placer/actividad" → -ING}
\begin{itemize}
    \item enjoy, love, hate, avoid, finish, suggest, recommend, consider
\end{itemize}

\textbf{💡 Truco: Verbos de "intención/deseo" → TO + infinitive}
\begin{itemize}
    \item want, hope, plan, decide, promise, refuse, agree, learn
\end{itemize}
\end{comparisonbox}

\section{Key Concepts: Verb Patterns}

In English, when one verb follows another, the second verb changes its form.

\begin{vocabbox}[title=Pattern 1: Verb + -ing (Gerund)]
Use \textbf{-ing} after these verbs:
\begin{itemize}
    \item \textbf{Enjoy:} I enjoy \textbf{reading}. \trans{Disfrutar}
    \item \textbf{Avoid:} She avoids \textbf{driving} at night. \trans{Evitar}
    \item \textbf{Finish:} He finished \textbf{working}. \trans{Terminar}
    \item \textbf{Suggest:} I suggest \textbf{going} by bus. \trans{Sugerir}
    \item \textbf{Recommend:} They recommend \textbf{booking} early. \trans{Recomendar}
    \item \textbf{Consider:} We considered \textbf{moving}. \trans{Considerar}
\end{itemize}
\end{vocabbox}

\begin{vocabbox}[title=Pattern 2: Verb + to + Infinitive]
Use \textbf{to + verb} after these verbs:
\begin{itemize}
    \item \textbf{Want:} I want \textbf{to sleep}. \trans{Querer}
    \item \textbf{Hope:} I hope \textbf{to see} you. \trans{Esperar (deseo)}
    \item \textbf{Decide:} She decided \textbf{to leave}. \trans{Decidir}
    \item \textbf{Plan:} We plan \textbf{to visit}. \trans{Planear}
    \item \textbf{Promise:} He promised \textbf{to help}. \trans{Prometer}
    \item \textbf{Refuse:} They refused \textbf{to pay}. \trans{Rehusar}
    \item \textbf{Learn:} I learned \textbf{to swim}. \trans{Aprender}
\end{itemize}
\end{vocabbox}

\section{Grammar Focus: Common Mistakes}

\begin{grammarbox}[title=Watch Out!]
\begin{itemize}
    \item \textbf{Incorrect:} I enjoy to play football.
    \item \textbf{Correct:} I enjoy \textbf{playing} football.
    \item \textbf{Incorrect:} I want going home.
    \item \textbf{Correct:} I want \textbf{to go} home.
\end{itemize}
\end{grammarbox}

\section{Practice Exercises}

\subsection{Exercise 1: Gerund or Infinitive?}
Choose the correct form.

\begin{enumerate}
    \item I promise (helping / to help) you tomorrow.
    \item She suggested (going / to go) to the park.
    \item We decided (buying / to buy) a new car.
    \item He finished (eating / to eat) his dinner.
    \item Do you enjoy (watching / to watch) TV?
\end{enumerate}

\subsection{Exercise 2: Categorize the Verbs}
Put these verbs in the correct column: \textit{avoid, learn, offer, consider, expect, mind}.

\begin{table}[h]
\centering
\begin{tabular}{|c|c|}
\hline
\textbf{+ -ING} & \textbf{+ TO + VERB} \\
\hline
\hspace{4cm} & \hspace{4cm} \\
\hspace{4cm} & \hspace{4cm} \\
\hspace{4cm} & \hspace{4cm} \\
\hline
\end{tabular}
\end{table}

\subsection{Exercise 3: Correct the Text}
Find and correct 5 mistakes in this text about Sharon.

\textit{Sharon is a teacher. She enjoys to teach and meet new people. In her free time, she likes watch documentaries. She plans visiting Ireland next year. She wants learning French.}

\begin{tcolorbox}[colback=white,height=4cm]
% Write corrected text here...
\end{tcolorbox}

\subsection{Exercise 4: Writing Task}
Write sentences about yourself using:
\begin{itemize}
    \item I plan...
    \item I avoid...
    \item I enjoy...
    \item I hope...
\end{itemize}

\begin{tcolorbox}[colback=white,height=5cm]
% Write here...
\end{tcolorbox}

\section{Key Takeaways}
\begin{itemize}
    \item Memorize which verbs take \textbf{-ing} (enjoy, avoid, suggest).
    \item Memorize which verbs take \textbf{to + infinitive} (want, hope, decide).
    \item \textbf{Recommend} and \textbf{Suggest} are followed by -ing (I recommend visiting...).
\end{itemize}

\section{Online Practice}
\begin{tcolorbox}[colback=blue!5,colframe=blue!40!black,title=Resources to Practice]
Online PracticeReinforce what you have learned with these interactive exercises:
\begin{itemize}
    \item \textbf{Gerunds \& Infinitives:} \url{https://www.perfect-english-grammar.com/gerunds-and-infinitives-exercise-1.html}
    \item \textbf{Verb Patterns Practice:} \url{https://test-english.com/grammar-points/b1-b2/verb-patterns-gerund-infinitive/}
    \item \textbf{Interactive Exercises:} \url{https://learnenglish.britishcouncil.org/grammar/intermediate-to-upper-intermediate/verbs-followed-by-ing-or-infinitive}
    \item \textbf{Complete List:} \url{https://www.englishpage.com/gerunds/gerund_infinitive_list.htm}
\end{itemize}
\end{tcolorbox}
