\chapter{Travel, Customs, and Superstitions}

\begin{center}
\begin{tabular}{ccc}
\cefrlevel{B2-C1} & \textbf{Study Time:} 3-4 hours & \textbf{Difficulty:} ⭐⭐⭐⭐☆
\end{tabular}
\end{center}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item Vocabulary related to travel and tourism
    \item How different cultures have different customs
    \item Common superstitions across cultures
    \item How to discuss cultural differences respectfully
    \item Using English to describe travel experiences
\end{itemize}

\section{Reading Context}

\begin{readingbox}[title=Dialogue: Discussing Cultural Customs]
\textbf{Maria:} I'm visiting Spain next month. I'm excited but also nervous about cultural differences.\\
\textbf{James:} Don't worry! I visited Spain last year. One custom I learned is that dinner is much later than in the UK.\\
\textbf{Maria:} Really? How much later?\\
\textbf{James:} Usually around 9 PM. Also, in some regions, there are superstitions about certain numbers and colors.\\
\textbf{Maria:} Interesting! What about greetings?\\
\textbf{James:} Spaniards are very warm and often greet with two kisses on the cheek. It's a sign of friendliness.\\
\textbf{Maria:} That's helpful to know. I want to be respectful of their traditions.
\end{readingbox}

\section{Key Vocabulary}

\begin{vocabbox}[title=Travel and Cultural Terms]
\begin{itemize}
    \item \keyterm{Custom} - Traditional practice or way of behaving
    \item \keyterm{Tradition} - Long-established practice passed down
    \item \keyterm{Superstition} - Belief not based on reason or fact
    \item \keyterm{Etiquette} - Rules for polite behavior in society
    \item \keyterm{Heritage} - Property, tradition, or culture inherited
    \item \keyterm{Festival} - Celebration or event
    \item \keyterm{Ritual} - Religious or solemn ceremony
\end{itemize}
\end{vocabbox}

\section{Customs Around the World}

\subsection{Dining Customs}

\begin{table}[h]
\centering
\begin{tabular}{|l|p{5cm}|}
\hline
\textbf{Country/Region} & \textbf{Custom} \\
\hline
Japan & Remove shoes before entering homes; slurp noodles \\
\hline
India & Use right hand for eating (left hand considered impolite) \\
\hline
Spain & Dinner is late (8-9 PM); long meals are social \\
\hline
France & Bread is never cut with a knife; bread is broken by hand \\
\hline
Middle East & Accept tea or coffee as a sign of hospitality \\
\hline
\end{tabular}
\caption{Dining customs by region}
\end{table}

\subsection{Greeting Customs}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Region} & \textbf{Greeting} \\
\hline
Spain & Two cheek kisses \\
\hline
France & Two or three cheek kisses depending on region \\
\hline
Japan & Bow (depth depends on status/respect) \\
\hline
Middle East & Hand to heart gesture with greeting \\
\hline
Thailand & Wai (hands pressed together) \\
\hline
UK & Handshake or verbal greeting \\
\hline
\end{tabular}
\caption{Greetings around the world}
\end{table}

\section{Common Superstitions}

\begin{vocabbox}[title=Superstitions Across Cultures]
\begin{itemize}
    \item \textbf{Number 13:} Considered unlucky in Western cultures
    \item \textbf{Black cat:} Bad luck if it crosses your path (Western belief)
    \item \textbf{Red:} Prosperity and good luck in Chinese culture
    \item \textbf{Broken mirror:} Seven years bad luck (Western)
    \item \textbf{Knocking on wood:} Protection from bad luck
    \item \textbf{Finding a penny:} Good luck if heads up
\end{itemize}
\end{vocabbox}

\section{Practice Exercises}

\subsection{Exercise 1: Discuss Cultural Customs}
Choose a country you'd like to visit. Research and describe:
\begin{itemize}
    \item Important customs
    \item Greeting etiquette
    \item Dining customs
    \item Any superstitions
\end{itemize}

\section{Key Takeaways}

\begin{itemize}
    \item Understanding customs shows respect for other cultures
    \item Customs vary significantly across regions
    \item Superstitions reflect cultural beliefs and values
    \item Always ask respectfully about unfamiliar customs
    \item Being culturally aware improves travel experiences
\end{itemize}


\section{Online Practice}
Here are some useful websites to practice this topic:
\begin{itemize}
    \item \href{https://learnenglish.britishcouncil.org/grammar}{British Council LearnEnglish}
    \item \href{https://test-english.com/}{Test-English}
    \item \href{https://www.perfect-english-grammar.com/}{Perfect English Grammar}
\end{itemize}
