\chapter{Email and Written Business Communication}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item How to structure a professional email
    \item Appropriate tone and register for business writing
    \item Key phrases for opening, closing, and various business situations
    \item How to write formal and informal emails
    \item Common conventions in British business communication
    \item How to make emails clear, concise, and professional
\end{itemize}

\section{Reading Context: Email Examples}

\begin{readingbox}[title=Example 1: Professional Inquiry Email]
\textbf{Subject:} Project Timeline - Request for Meeting

Dear Mr. Johnson,

I \textbf{hope this email finds you well}. I am writing to inquire about the possibility of scheduling a meeting to discuss the upcoming project timeline.

I \textbf{would appreciate it if} we could meet next week to review the deliverables and confirm the deadline. Would you be available on Tuesday or Wednesday afternoon?

I \textbf{look forward to hearing from you}.

Best regards,\\
Sarah Mitchell\\
Project Coordinator
\end{readingbox}

\begin{readingbox}[title=Example 2: Informal Email to Colleague]
Hi Tom,

Quick heads-up — the client has pushed back the presentation to next Friday. I've updated the schedule on the shared drive.

Let me know if you need any changes to the agenda. I \textbf{wanted to touch base} before the meeting to align on the key points.

Cheers,\\
Emma
\end{readingbox}

\section{Grammar Focus: Email Structure and Tone}

\begin{grammarbox}[title=Formal Email Structure]
\textbf{1. Subject Line:}
\begin{itemize}
    \item Should be specific and clear
    \item Examples: "Request for Meeting," "Project Update - Q1 Progress"
\end{itemize}

\textbf{2. Greeting:}
\begin{itemize}
    \item Formal: "Dear Mr./Ms. [Last Name]," "Dear Sir/Madam,"
    \item Less formal: "Hello [Name]," "Hi [Name],"
\end{itemize}

\textbf{3. Opening:}
\begin{itemize}
    \item Formal: "I hope this email finds you well."
    \item "I am writing to..." (state purpose)
\end{itemize}

\textbf{4. Body:}
\begin{itemize}
    \item Clear paragraphs with one main idea each
    \item Use active voice when possible
    \item Be concise and professional
\end{itemize}

\textbf{5. Closing:}
\begin{itemize}
    \item Formal: "Best regards," "Yours sincerely," "Respectfully,"
    \item Less formal: "Best," "Thanks," "Cheers,"
\end{itemize}

\textbf{6. Signature:}
\begin{itemize}
    \item Full name
    \item Title
    \item Contact information
\end{itemize}
\end{grammarbox}

\section{Key Phrases for Business Emails}

\begin{vocabbox}[title=Essential Business Email Phrases]

\textbf{Opening Statements:}
\begin{itemize}
    \item I \textbf{hope this email finds you well}.
    \item Thank you for your \textbf{prompt response}.
    \item I \textbf{wanted to follow up on} our previous conversation.
    \item I am \textbf{writing to inform you that}...
    \item I \textbf{am reaching out} to discuss...
\end{itemize}

\textbf{Making Requests:}
\begin{itemize}
    \item I \textbf{would appreciate it if} you could...
    \item \textbf{Could you please}... \trans{¿Podrías por favor...?}
    \item \textbf{Would you mind}... \trans{¿Te importaría...?}
    \item I \textbf{would be grateful if} you could...
    \item \textbf{At your earliest convenience}, please...
\end{itemize}

\textbf{Providing Information:}
\begin{itemize}
    \item \textbf{I would like to inform you that}...
    \item Please \textbf{find attached} the document.
    \item \textbf{As discussed}, we will proceed with...
    \item \textbf{In light of} the recent changes...
\end{itemize}

\textbf{Closing Statements:}
\begin{itemize}
    \item I \textbf{look forward to hearing from you}.
    \item Please \textbf{let me know if} you have any questions.
    \item \textbf{Should you need any further information}, do not hesitate to contact me.
    \item Thank you for your \textbf{attention to this matter}.
\end{itemize}

\textbf{Casual/Informal:}
\begin{itemize}
    \item \textbf{Just to touch base}...
    \item \textbf{Quick heads-up}...
    \item \textbf{Thanks for the heads up}.
    \item \textbf{Let me know} if you need anything.
    \item \textbf{Cheers} (British informal goodbye)
\end{itemize}

\end{vocabbox}

\section{Types of Business Emails}

\subsection{1. Request for Information}

\textbf{Pattern:}
\begin{itemize}
    \item Opening: State the purpose
    \item Body: Explain what you need and why
    \item Closing: Request timeline and express gratitude
\end{itemize}

\textbf{Example:}

"Dear Ms. Brown, I would appreciate it if you could provide information regarding the updated project specifications. Could you send them by Friday? Thank you in advance."

\subsection{2. Status Update}

\textbf{Pattern:}
\begin{itemize}
    \item Opening: State that it's an update
    \item Body: Current progress, achievements, next steps
    \item Closing: Note any pending items or decisions needed
\end{itemize}

\textbf{Example:}

"Hi James, I'm writing to update you on the marketing campaign. We've completed the first draft of the advertisement, and it's ready for your review. Please let me know your feedback by Wednesday."

\subsection{3. Complaint or Problem Resolution}

\textbf{Pattern:}
\begin{itemize}
    \item Opening: Acknowledge the relationship
    \item Body: Describe the issue professionally (no blame)
    \item Closing: Propose solutions and request action
\end{itemize}

\textbf{Example:}

"Dear Mr. Chen, Thank you for your business. I wanted to bring a concern to your attention regarding the last shipment. The items arrived damaged. Could we arrange for a replacement? Thank you for your prompt attention to this matter."

\section{Tone and Register in Business Writing}

\begin{table}[h]
\centering
\begin{tabular}{|l|p{4cm}|p{4cm}|}
\hline
\textbf{Context} & \textbf{Register} & \textbf{Example Phrase} \\
\hline
To senior management & Very Formal & "I would humbly request..." \\
\hline
To a colleague & Semi-formal & "I wanted to check if..." \\
\hline
To a team member & Informal & "Quick question..." \\
\hline
To a new client & Formal & "I am delighted to..." \\
\hline
Regular correspondence & Semi-formal & "Thank you for..." \\
\hline
\end{tabular}
\caption{Email register by context}
\end{table}

\section{Practice Exercises}

\subsection{Exercise 1: Complete the Email}
Fill in the blanks with appropriate phrases:

Dear Mr. Harrison,

\underline{\hspace{4cm}}, I am writing to discuss our Q2 strategy. I \underline{\hspace{4cm}} if we could schedule a meeting next week to review the projections.

\underline{\hspace{4cm}}, I have attached the preliminary report. \underline{\hspace{4cm}} your feedback before proceeding further.

\underline{\hspace{4cm}}.

Best regards,

\subsection{Exercise 2: Identify Formal vs. Informal}
Label each phrase as F (Formal) or I (Informal):

\begin{enumerate}
    \item I hope this email finds you well. \underline{\hspace{2cm}}
    \item Just a quick heads-up. \underline{\hspace{2cm}}
    \item I would appreciate it if you could. \underline{\hspace{2cm}}
    \item Let me know ASAP. \underline{\hspace{2cm}}
\end{enumerate}

\subsection{Exercise 3: Write Your Own Professional Email}
Write an email requesting a meeting with your supervisor about a project. Include:
\begin{itemize}
    \item Appropriate subject line
    \item Formal greeting
    \item Clear purpose statement
    \item Specific request with timeline
    \item Professional closing
\end{itemize}

\begin{tcolorbox}[colback=white,height=8cm]
\end{tcolorbox}

\section{[INCORRECT] Common Mistakes in Business Email}

\begin{tcolorbox}[colback=red!5, colframe=red!30, title=Email Etiquette Errors]

\textbf{Mistake 1:} Too informal tone with higher authority
\begin{itemize}
    \item [INCORRECT] "Hey boss, whatcha up to?"
    \item [CORRECT] "Dear Sir/Madam, I hope this email finds you well."
\end{itemize}

\textbf{Mistake 2:} Missing subject line or vague subject
\begin{itemize}
    \item [INCORRECT] Subject: "Hi"
    \item [CORRECT] Subject: "Request for Project Extension - Deadline Update"
\end{itemize}

\textbf{Mistake 3:} Unclear opening
\begin{itemize}
    \item [INCORRECT] "I'm writing about something."
    \item [CORRECT] "I am writing to request approval for the budget proposal."
\end{itemize}

\textbf{Mistake 4:} Forgetting attachments or mentioning them incorrectly
\begin{itemize}
    \item [INCORRECT] "Please find the attached document" (but file is missing)
    \item [CORRECT] Make sure to actually attach before sending
\end{itemize}

\textbf{Mistake 5:} Poor grammar or spelling
\begin{itemize}
    \item [INCORRECT] "Their budget needs approval"
    \item [CORRECT] "The budget needs approval" or "Their approval is needed"
\end{itemize}

\end{tcolorbox}

\section{🇬🇧 British Business Email Conventions}

British business emails have distinct characteristics:

\begin{itemize}
    \item \textbf{Politeness markers:} British emails are often more formal and polite
    \begin{itemize}
        \item "I would be grateful if..." (more British than American)
        \item "Yours faithfully," (traditional British closing)
    \end{itemize}
    \item \textbf{Spelling:} Use British English spellings
    \begin{itemize}
        \item colour, organisation, recognise
    \end{itemize}
    \item \textbf{Tone:} British writers often use understated language
    \begin{itemize}
        \item "This might be worth considering..." (rather than direct statements)
    \end{itemize}
    \item \textbf{Closings:} Common British sign-offs
    \begin{itemize}
        \item "Kind regards," "Yours sincerely," "Cheers," "All the best"
    \end{itemize}
\end{itemize}

\section{Key Takeaways}

\begin{itemize}
    \item Always include a clear subject line
    \item Match your tone (formal/informal) to the recipient and context
    \item Use standard structure: greeting, opening, body, closing, signature
    \item Use appropriate phrases like "I would appreciate," "Could you please," "Look forward to"
    \item Proofread before sending for grammar and spelling errors
    \item Be concise and clear — respect the reader's time
    \item Use "I would" and "could you" for polite requests
    \item Always express gratitude appropriately
    \item Follow British conventions if writing to UK audiences
\end{itemize}
