\chapter{Present Perfect Continuous}

\begin{center}
\begin{tabular}{ccc}
\cefrlevel{B1-B2} & \textbf{Study Time:} 3-4 hours & \textbf{Difficulty:} ⭐⭐⭐⭐☆
\end{tabular}
\end{center}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item How to form the Present Perfect Continuous tense
    \item When to use Present Perfect Continuous vs. Present Perfect Simple
    \item The difference between "for" and "since"
    \item How to talk about duration and ongoing actions
\end{itemize}

\section{Reading Context}
\begin{readingbox}[title=Dialogue: Catching Up]
\textbf{Tom:} Hi Sarah! I haven't seen you for ages. What \textbf{have you been doing}?\\
\textbf{Sarah:} Hey Tom! I\textbf{'ve been working} a lot. I started a new job in January.\\
\textbf{Tom:} Really? How long \textbf{have you been working} there?\\
\textbf{Sarah:} For about six months. What about you?\\
\textbf{Tom:} I\textbf{'ve been traveling}. I just got back from Asia.\\
\textbf{Sarah:} Wow! \textbf{Have you been learning} any new languages?\\
\textbf{Tom:} A little bit. I\textbf{'ve been trying} to learn Japanese \textbf{since} I visited Tokyo.
\end{readingbox}

\begin{comparisonbox}[title=🇪🇸 Present Perfect Continuous vs Simple: Español vs Inglés]
\textbf{⚠️ En español NO distinguimos proceso vs resultado como en inglés}

\textbf{1. Español - un solo tiempo:}
\begin{itemize}
    \item "He estado estudiando 3 horas" (proceso y duración)
    \item "He estudiado 3 capítulos" (resultado y cantidad)
    \item Ambas usan el mismo tiempo verbal en español
\end{itemize}

\textbf{2. Inglés - dos tiempos diferentes:}
\begin{itemize}
    \item "I\textbf{'ve been studying} for 3 hours" (Continuous - proceso, duración)
    \item "I\textbf{'ve studied} 3 chapters" (Simple - resultado, cantidad completada)
\end{itemize}

\textbf{3. Regla clave - ¿Proceso o Resultado?}
\begin{center}
\textbf{Continuous} → Énfasis en CUÁNTO TIEMPO (for/since)\\
\textbf{Simple} → Énfasis en CUÁNTO HAS HECHO (cantidad, completado)
\end{center}

\textbf{4. Ejemplos contrastivos:}
\begin{itemize}
    \item "He estado leyendo" = I\textbf{'ve been reading} (proceso) ✓
    \item "He leído 5 libros" = I\textbf{'ve read} 5 books (resultado) ✓
    \item "Vivo aquí desde 2020" = I\textbf{'ve been living} here since 2020 (continuo)
    \item "He vivido en 3 países" = I\textbf{'ve lived} in 3 countries (experiencia)
\end{itemize}

\textbf{💡 Para hispanohablantes:}
\begin{itemize}
    \item Si dices "CUÁNTO TIEMPO" → Continuous
    \item Si dices "CUÁNTAS VECES/CANTIDAD" → Simple
\end{itemize}
\end{comparisonbox}

\section{Grammar Focus: Present Perfect Continuous}

This tense is used for actions that started in the past and continue until now, or have recently stopped with visible results.

\begin{grammarbox}[title=Structure]
\begin{center}
\Large Subject + \textbf{have/has been} + \textbf{verb-ing}
\end{center}

\textbf{Examples:}
\begin{itemize}
    \item I \textbf{have been studying} English for 3 years. \trans{He estado estudiando...}
    \item She \textbf{has been waiting} since 9 AM. \trans{Ella ha estado esperando...}
\end{itemize}
\end{grammarbox}

\section{Key Concepts: For vs. Since}

\begin{vocabbox}[title=Time Expressions]
\textbf{FOR} + Period of Time (Duration)
\begin{itemize}
    \item for 2 hours \trans{por 2 horas}
    \item for a long time \trans{por mucho tiempo}
    \item for 10 years \trans{por 10 años}
\end{itemize}

\textbf{SINCE} + Point in Time (Start date)
\begin{itemize}
    \item since Monday \trans{desde el lunes}
    \item since 2010 \trans{desde 2010}
    \item since I was a child \trans{desde que era niño}
\end{itemize}
\end{vocabbox}

\section{Present Perfect Continuous vs. Simple}

\begin{table}[h]
\centering
\begin{tabular}{|p{6cm}|p{6cm}|}
\hline
\textbf{Continuous (Process/Duration)} & \textbf{Simple (Result/Completion)} \\
\hline
I \textbf{have been reading} this book. (Still reading) & I \textbf{have read} this book. (Finished) \\
\hline
She \textbf{has been cooking} all day. (Activity) & She \textbf{has cooked} dinner. (Done) \\
\hline
\end{tabular}
\caption{Continuous vs. Simple}
\end{table}

\section{Practice Exercises}

\subsection{Exercise 1: For or Since?}
Complete with "for" or "since":

\begin{enumerate}
    \item I've been waiting \underline{\hspace{2cm}} 20 minutes.
    \item She's been living in London \underline{\hspace{2cm}} 2019.
    \item We've been married \underline{\hspace{2cm}} 10 years.
    \item He's been sleeping \underline{\hspace{2cm}} this morning.
\end{enumerate}

\subsection{Exercise 2: Correct the Errors}
Find and correct the mistakes:

\begin{enumerate}
    \item I have been learning English since 6 years. $\rightarrow$ \underline{\hspace{5cm}}
    \item I am living here since 2020. $\rightarrow$ \underline{\hspace{5cm}}
    \item How long you have been waiting? $\rightarrow$ \underline{\hspace{5cm}}
\end{enumerate}

\subsection{Exercise 3: Personal Questions}
Answer these questions about yourself:

\begin{enumerate}
    \item How long have you been studying English?
    \item How long have you been living in your current city?
\end{enumerate}

\begin{tcolorbox}[colback=white,height=4cm]
% Write your answers here
\end{tcolorbox}

\section{Key Takeaways}
\begin{itemize}
    \item Use Present Perfect Continuous for ongoing actions (have been -ing).
    \item Use "for" with a duration (for 2 years).
    \item Use "since" with a starting point (since 2020).
    \item Stative verbs (know, like, believe) are usually NOT used in continuous forms.
\end{itemize}

\section{Online Practice}
\begin{tcolorbox}[colback=blue!5,colframe=blue!40!black,title=Resources to Practice]
Online PracticeReinforce what you have learned with these interactive exercises:
\begin{itemize}
    \item \textbf{Present Perfect Continuous:} \url{https://www.perfect-english-grammar.com/present-perfect-continuous-exercise-1.html}
    \item \textbf{For vs Since:} \url{https://test-english.com/grammar-points/a2/for-since/}
    \item \textbf{Interactive Practice:} \url{https://learnenglish.britishcouncil.org/grammar/english-grammar-reference/present-perfect-continuous}
    \item \textbf{Stative Verbs:} \url{https://www.englishpage.com/verbpage/presentperfectcontinuous.html}
\end{itemize}
\end{tcolorbox}