
\chapter{Vocabulary}

This chapter contains useful vocabulary words and phrases that were covered in previous classes. It includes definitions and example sentences to help you understand how to use them in context.

\section{Vocabulary}

\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Vocabulary} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
bursary & a scholarship to attend a college or university. & She received a bursary to help pay for her tuition. \\
\hline
career break & a period of time when someone stops working in their profession to rest or pursue other interests. & She took a career break to travel the world. \\
\hline
coach & a bus used for long-distance travel. & We took a coach to the countryside for the weekend. \\
\hline
commute & travel some distance between one's home and place of work on a regular basis. & I commute to the city every day for work. \\
\hline
crutches & a support used by a person who is unable to walk without assistance. & He had to use crutches after breaking his leg. \\
\hline
delightful & highly pleasing. & The cake she baked was absolutely delightful. \\
\hline
gap year & a year spent away from formal education, typically to travel or gain work experience. & He took a gap year before starting university. \\
\hline
get up to & to do something, often something bad. & What did you get up to at the party last night? \\
\hline
hangover & a painful physical condition following the excessive consumption of alcohol. & He had a terrible hangover after the party. \\
\hline
milestone & a significant event or stage in the development of something. & Graduating from college is a major milestone in life. \\
\hline
mishap & an unfortunate accident. & He had a mishap while hiking and sprained his ankle. \\
\hline
nicked & stolen (informal). & My car was nicked last night. \\
\hline
nightshifts & working during the night hours. & She works the nightshifts at the hospital. \\
\hline
outstanding & exceptionally good. & Her performance in the play was outstanding. \\
\hline
perk & a benefit or advantage. & One of the perks of the job is free lunch. \\
\hline
placement & the action of placing someone in a job or position. & The placement of the new employee was successful. \\
\hline
proofread & read (printer's proofs or other written or printed material) and mark any errors. & I need to proofread my essay before submitting it. \\
\hline
purse & a small bag used by women to carry money. & I am looking for my purse. \\
\hline
revision & the action of revising. & The document is under revision for accuracy. \\
\hline
sandal & a light shoe with either an openwork upper or straps attaching the sole to the foot. & She wore sandals to the beach. \\
\hline
scholarship & a grant or payment made to support a student's education, awarded on the basis of academic or other achievement. & She received a scholarship for her excellent grades. \\
\hline
shy & being reserved or having or showing nervousness or timidity in the company of other people. & She is shy and doesn't like speaking in public. \\
\hline
sightseeing & visiting famous places. & I went sightseeing in London. \\
\hline
stomach ache & dolor de estómago. & I have a stomach ache. \\
\hline
stuck & unable to move or be moved. & The car got stuck in the mud. \\
\hline
thin & having little, or too little, flesh or fat on the body. & The thin cat looked hungry. \\
\hline
mob & a large crowd of people, especially one that is disorderly and intent on causing trouble or violence. & A mob gathered outside the courthouse. \\
\hline
mobbing & the act of a large group of people attacking or overwhelming someone. & The celebrity was mobbed by fans at the airport. \\
\hline
rigorous & extremely thorough and careful. & The rigorous training prepared the athletes for the competition. \\
\hline
thorough & complete with regard to every detail; not superficial or partial. & The detective conducted a thorough investigation. \\
\hline
remains & the parts left over after other parts have been removed, used, or destroyed. & The remains of the ancient building were discovered by archaeologists. \\
\hline
remnants & a small remaining quantity of something. & The remnants of the ancient building were discovered by archaeologists. \\
\hline
leftovers & food remaining after a meal. & We had leftovers from dinner last night. \\
\hline
vague & of uncertain, indefinite, or unclear character or meaning. & His vague answers made it hard to understand his intentions. \\
\hline
wearily & in a tired or exhausted manner. & She spoke wearily after a long day at work. \\
\hline
innate & inborn; natural. & She has an innate talent for music. \\
\hline
upbringing & the treatment and instruction received by a child from its parents throughout its childhood. & Her strict upbringing influenced her disciplined lifestyle. \\
\hline
lie-in & the act of staying in bed later than usual in the morning. & On weekends, I like to have a lie-in. \\
\hline
networking & the action or process of interacting with others to exchange information and develop professional or social contacts. & Networking is important for career growth. \\
\hline
flee & run away from a place or situation of danger. & The villagers had to flee during the war. \\
\hline
affordability & the quality of being affordable; inexpensive. & The affordability of housing is a major concern in the city. \\
\hline
sprawling & spreading out over a large area in an untidy or irregular way. & The sprawling city has many suburbs. \\
\hline
\end{longtable}


\section{Informal Contractions}
\begin{longtable}{lll}
\textbf{Contraction} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
gonna & going to & I'm gonna study tonight \\
wanna & want to & Do you wanna go? \\
gotta & got to & I gotta leave now \\
kinda & kind of & It's kinda cold today \\
gimme & give me & Gimme a break \\
cya & see you later & Cya tomorrow \\
\end{longtable}
    
\section{British Slang \& Expressions}
\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Expression} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
a bloke & a man & There's a bloke at the door \\
a catch up & meet with someone to talk & We should have a catch up \\
a tenner & a 10 pound note & Can you lend me a tenner? \\
cheap as chips & very cheap & That shirt was cheap as chips \\
chop chop & hurry up & Come on, chop chop \\
dodgy & untrustworthy, suspicious & That looks a bit dodgy \\
fancy & I really want to & I fancy a coffee \\
feel right as rain & feel well again & I feel right as rain now \\
give someone a bell & call someone & Give me a bell later \\
grab some food & get something to eat & Let's grab some food \\
gutted & very disappointed & I was gutted to miss it \\
-ish & approximately & Meet me at 3-ish \\
knackered & very tired & I'm absolutely knackered \\
on and off & not consistently, not always & It's been raining on and off \\
peckish & snack & I'm feeling a bit peckish \\
shuteye & sleep & I need some shuteye \\
skint & broke, no money temporarily & I'm a bit skint this week \\
up for something & agree to do something & Are you up for a movie? \\
went off & exploded, made a loud noise & The alarm went off at 6am \\

\end{longtable}

\section{General Expressions}
\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Expression} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
bare-feet & without shoes & He walked barefoot on the grass \\
encourage & give support, confidence, or hope to (someone) & She encouraged me to apply \\
fire away & is used to give someone permission to ask questions or make comments & Fire away, I'm ready \\
get back in touch & contact someone again & I'll get back in touch next week \\
hop & jump on one foot & The kids like to hop around \\
pay attention & focus on something & Please pay attention to the instructions \\
ran me over & hit me with a vehicle & A car ran me over \\
ran out of & have no more of (something) & I've run out of time \\
stay out & remain outside & We decided to stay out late \\
take it easy & relax & Take it easy this weekend \\
wander around & walk around without a specific purpose & We wandered around the city \\
\end{longtable}


\section{Order Words/Phrases}
\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Word/Phrase} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
after that & following that; subsequently & After that, we can make a decision \\
\hline
finally & at the end; lastly & Finally, we will summarize our findings \\
\hline
firstly & used to introduce the first point or reason in a list & Firstly, we need to gather information \\
\hline
following that & after that; subsequently & Following that, we will review the plan \\
\hline
next & after that; then & Next, we will discuss the results \\
\hline
secondly & used to introduce the second point or reason in a list & Secondly, we need to analyze the data \\
\hline


\end{longtable} 

\section{Collocations}
\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Preposition} & \textbf{Usage} & \textbf{Example} \\
\hline
\endhead
at & used with "the weekend" & At the weekend \\
\hline
in & used with parts of the day & In the morning/afternoon/evening \\
\hline
on & used with days & On Saturday \\
\hline
\end{longtable} 



\section{Phrasal Verbs}
\begin{longtable}{p{0.25\textwidth}p{0.35\textwidth}p{0.35\textwidth}}
\textbf{Phrasal Verb} & \textbf{Meaning} & \textbf{Example} \\
\hline
\endhead
break down & to stop functioning (for machines) & My car broke down on the way to work \\
\hline
call off & to cancel something & They called off the meeting \\
\hline
carry on & to continue doing something & She carried on working despite the noise \\
\hline
check in & to register at a hotel or airport & We checked in at the hotel at noon \\
\hline
check out & to leave a hotel after paying the bill & We checked out of the hotel in the morning \\
\hline
fill out & to complete a form & Please fill out this application form \\
\hline
give up & to stop trying or doing something & He gave up smoking last year \\
\hline
look after & to take care of someone or something & She looks after her younger brother \\
\hline
look for & to search for something & I'm looking for my keys \\
\hline
pick up & to collect someone or something & I'll pick you up at the airport \\
\hline
set up & to arrange or organize something & They set up a new company \\
\hline
turn down & to refuse or reject something & She turned down the job offer \\
\hline
turn up & to arrive or appear somewhere & He turned up late to the meeting \\
\hline
cut down on & to reduce the amount of something & I'm trying to cut down on sugar \\
\hline
work out & to exercise or train physically & I work out at the gym every day \\
\hline
pick up & to start learning or doing something new & She picked up Spanish quickly \\
\hline
save up & to accumulate money for a specific purpose & I'm saving up for a new car \\
\hline
spend more time & to allocate time for a particular activity or person & I want to spend more time with my family \\
\hline
make up & to reconcile after a disagreement & They made up after their argument \\
\hline
take up & to begin a new activity or hobby & He took up painting last year \\
\hline
plan out & to organize or arrange something in detail & We need to plan out our trip \\
\hline
follow through & to complete or carry out a plan or goal & She followed through on her promise \\
\hline
slow down & to reduce speed or take things more leisurely & You should slow down and relax \\
\hline
carry out & to complete something & We need to carry out the plan \\
\hline
cut down on & to reduce & I'm trying to cut down on sugar \\
\hline
get round to & to find time to do something & I'll get round to it when I can \\
\hline
pay off & to bring good results & Your hard work will pay off \\
\hline
map out & to plan in detail & We need to map out our strategy \\
\hline
take up & to start something & She took up painting last year \\
\hline
stick to & to continue with something & I'm sticking to my diet \\
\hline
follow through on & to finish what you started & She followed through on her promise \\
\hline 
\end{longtable}




