\chapter{Reported Speech}

\section{Lesson Objectives}
In this chapter, you will learn:
\begin{itemize}
    \item How to report what someone said (statements)
    \item How to report questions
    \item How to report commands and requests
    \item Tense changes in reported speech (backshift)
    \item Time and place expression changes
    \item Reporting verbs (say, tell, ask, etc.)
\end{itemize}

\section{Reading Context}
\begin{readingbox}[title=Dialogue: Reporting a Conversation]
\textbf{Anna:} Did you talk to the manager?\\
\textbf{Ben:} Yes. He \textbf{said} that he \textbf{was} very busy.\\
\textbf{Anna:} What else did he say?\\
\textbf{Ben:} He \textbf{told} me that the meeting \textbf{had been cancelled}.\\
\textbf{Anna:} Oh no! Did he say why?\\
\textbf{Ben:} He \textbf{explained} that he \textbf{had to} go to head office. He also \textbf{asked} me if I \textbf{could} reschedule it.\\
\textbf{Anna:} What did you say?\\
\textbf{Ben:} I \textbf{told} him that I \textbf{would} check my calendar and \textbf{would let} him know.
\end{readingbox}

\section{Grammar Focus: What is Reported Speech?}

\textbf{Direct Speech:} Quoting someone's exact words with quotation marks.
\begin{itemize}
    \item Tom said, "I am tired."
\end{itemize}

\textbf{Reported Speech (Indirect Speech):} Reporting what someone said without quotation marks.
\begin{itemize}
    \item Tom said (that) he was tired.
\end{itemize}

\section{Reporting Statements}

\begin{grammarbox}[title=Basic Structure]
\textbf{Direct:} "I am happy."

\textbf{Reported:} He/She said (that) he/she was happy.

\textbf{Structure:} 
\begin{center}
\Large Subject + \textbf{said (that)} / \textbf{told me (that)} + reported clause
\end{center}

\textbf{Important Differences:}
\begin{itemize}
    \item \textbf{SAY:} He said (that)... (no object)
    \item \textbf{TELL:} He told me/her/them (that)... (needs object)
\end{itemize}
\end{grammarbox}

\section{Tense Changes (Backshift)}

When the reporting verb is in the past (said, told), we usually change the tenses backwards.

\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Direct Speech} & \textbf{Reported Speech} \\
\hline
Present Simple & Past Simple \\
"I work here." & She said she worked there. \\
\hline
Present Continuous & Past Continuous \\
"I am working." & She said she was working. \\
\hline
Past Simple & Past Perfect \\
"I worked yesterday." & She said she had worked the day before. \\
\hline
Present Perfect & Past Perfect \\
"I have worked here." & She said she had worked there. \\
\hline
Will & Would \\
"I will come." & She said she would come. \\
\hline
Can & Could \\
"I can help." & She said she could help. \\
\hline
May & Might \\
"I may go." & She said she might go. \\
\hline
Must & Had to \\
"I must leave." & She said she had to leave. \\
\hline
\end{tabular}
\caption{Tense changes in reported speech}
\end{table}

\section{Pronoun and Time/Place Changes}

\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Direct Speech} & \textbf{Reported Speech} \\
\hline
\multicolumn{2}{|c|}{\textbf{Pronouns}} \\
\hline
I / me & he/she / him/her \\
you & I / me (depending on context) \\
we / us & they / them \\
my & his/her \\
\hline
\multicolumn{2}{|c|}{\textbf{Time Expressions}} \\
\hline
today & that day \\
yesterday & the day before / the previous day \\
tomorrow & the next day / the following day \\
this week & that week \\
last week & the week before / the previous week \\
next week & the following week \\
ago & before \\
now & then / at that time \\
\hline
\multicolumn{2}{|c|}{\textbf{Place}} \\
\hline
here & there \\
this & that \\
these & those \\
\hline
\end{tabular}
\caption{Pronoun and expression changes}
\end{table}

\section{Examples of Reported Statements}

\begin{examplebox}[title=Statement Transformations]
\textbf{1. Present Simple $\rightarrow$ Past Simple}
\begin{itemize}
    \item Direct: "I \textbf{live} in London."
    \item Reported: She said (that) she \textbf{lived} in London.
\end{itemize}

\textbf{2. Present Continuous $\rightarrow$ Past Continuous}
\begin{itemize}
    \item Direct: "I \textbf{am studying} English."
    \item Reported: He said (that) he \textbf{was studying} English.
\end{itemize}

\textbf{3. Will $\rightarrow$ Would}
\begin{itemize}
    \item Direct: "I \textbf{will call} you tomorrow."
    \item Reported: She said (that) she \textbf{would call} me the next day.
\end{itemize}

\textbf{4. Past Simple $\rightarrow$ Past Perfect}
\begin{itemize}
    \item Direct: "I \textbf{went} to Paris last year."
    \item Reported: He said (that) he \textbf{had gone} to Paris the year before.
\end{itemize}
\end{examplebox}

\section{Reporting Questions}

Questions in reported speech don't use question word order or question marks.

\begin{grammarbox}[title=Yes/No Questions]
\textbf{Structure:} Subject + asked + \textbf{if/whether} + statement word order

\textbf{Direct:} "Do you like coffee?"

\textbf{Reported:} She asked (me) if/whether I liked coffee.

\textbf{More Examples:}
\begin{itemize}
    \item "Are you coming?" $\rightarrow$ He asked if I was coming.
    \item "Did you see her?" $\rightarrow$ She asked if I had seen her.
    \item "Can you help?" $\rightarrow$ He asked if I could help.
\end{itemize}
\end{grammarbox}

\begin{grammarbox}[title=WH-Questions]
\textbf{Structure:} Subject + asked + \textbf{WH-word} + statement word order

\textbf{Direct:} "Where do you live?"

\textbf{Reported:} She asked (me) where I lived.

\textbf{More Examples:}
\begin{itemize}
    \item "What are you doing?" $\rightarrow$ He asked what I was doing.
    \item "When did you arrive?" $\rightarrow$ She asked when I had arrived.
    \item "Why are you late?" $\rightarrow$ He asked why I was late.
\end{itemize}

\textbf{Important:} Use STATEMENT word order (subject + verb), NOT question order!
\end{grammarbox}

\section{Reporting Commands and Requests}

\begin{grammarbox}[title=Commands and Requests]
\textbf{Structure:} Subject + told/asked/ordered + object + \textbf{to} + infinitive

\textbf{Positive Commands:}
\begin{itemize}
    \item Direct: "Close the door."
    \item Reported: She \textbf{told} me \textbf{to close} the door.
\end{itemize}

\textbf{Negative Commands:}
\begin{itemize}
    \item Direct: "Don't be late."
    \item Reported: He \textbf{told} me \textbf{not to be} late.
\end{itemize}

\textbf{Requests (polite):}
\begin{itemize}
    \item Direct: "Could you help me, please?"
    \item Reported: She \textbf{asked} me \textbf{to help} her.
\end{itemize}
\end{grammarbox}

\section{Reporting Verbs}

\begin{vocabbox}[title=Common Reporting Verbs]
\begin{itemize}
    \item \textbf{say} (that): "I'm tired." $\rightarrow$ He said (that) he was tired.
    \item \textbf{tell} + object (that): "I'm busy." $\rightarrow$ She told me (that) she was busy.
    \item \textbf{ask} (if/whether): "Are you OK?" $\rightarrow$ He asked if I was OK.
    \item \textbf{ask} + WH: "Where do you live?" $\rightarrow$ She asked where I lived.
    \item \textbf{tell/ask} + to: "Close it." $\rightarrow$ He told me to close it.
    \item \textbf{explain} (that): "It's difficult." $\rightarrow$ She explained (that) it was difficult.
    \item \textbf{promise} (that): "I'll come." $\rightarrow$ He promised (that) he would come.
    \item \textbf{refuse} + to: "I won't do it." $\rightarrow$ She refused to do it.
    \item \textbf{offer} + to: "I'll help you." $\rightarrow$ He offered to help me.
    \item \textbf{suggest} + -ing: "Let's go." $\rightarrow$ She suggested going.
\end{itemize}
\end{vocabbox}

\section{When NOT to Backshift}

Sometimes we don't need to change the tense:

\begin{examplebox}[title=No Backshift Needed]
\textbf{1. When it's still true:}
\begin{itemize}
    \item "I live in London." $\rightarrow$ He said he lives in London. (still true)
\end{itemize}

\textbf{2. General truths/facts:}
\begin{itemize}
    \item "Water boils at 100°C." $\rightarrow$ She said water boils at 100°C.
\end{itemize}

\textbf{3. Recent speech (just now):}
\begin{itemize}
    \item "I'm tired." (just said) $\rightarrow$ She says she's tired.
\end{itemize}
\end{examplebox}

\section{Practice Exercises}

\subsection{Exercise 1: Change to Reported Speech}
Transform these direct statements.

\begin{enumerate}
    \item "I am happy." (She said...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
    \item "I work here." (He told me...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
    \item "I will call you tomorrow." (She said...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
    \item "I saw her yesterday." (He said...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
\end{enumerate}

\subsection{Exercise 2: Report These Questions}
\begin{enumerate}
    \item "Do you like coffee?" (She asked me...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
    \item "Where do you live?" (He asked me...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
    \item "Can you help me?" (She asked...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
    \item "When did you arrive?" (He asked...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
\end{enumerate}

\subsection{Exercise 3: Report Commands}
\begin{enumerate}
    \item "Close the window." (She told me...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
    \item "Don't be late." (He told me...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
    \item "Please help me." (She asked me...)\\
    $\rightarrow$ \underline{\hspace{8cm}}
\end{enumerate}

\subsection{Exercise 4: Say or Tell?}
Complete with the correct form of say or tell.

\begin{enumerate}
    \item He \underline{\hspace{3cm}} me that he was tired.
    \item She \underline{\hspace{3cm}} that she would come.
    \item They \underline{\hspace{3cm}} us to wait.
    \item I \underline{\hspace{3cm}} that I didn't know.
\end{enumerate}

\subsection{Exercise 5: Correct the Errors}
\begin{enumerate}
    \item He said me that he was busy.
    \item She asked where did I live.
    \item He told that he would come.
    \item She said me to close the door.
    \item He asked if was I ready.
\end{enumerate}

\subsection{Exercise 6: Complete Reported Speech}
Report this conversation:

\textbf{Tom:} "I'm very tired today."\\
\textbf{Sarah:} "Why? What did you do yesterday?"\\
\textbf{Tom:} "I worked late and didn't sleep well."\\
\textbf{Sarah:} "You should rest more."

\begin{tcolorbox}[colback=white,height=6cm]
Tom said...
Sarah asked...
Tom explained...
Sarah told him...
\end{tcolorbox}

\section{Key Takeaways}
\begin{itemize}
    \item \textbf{SAY:} He said (that)... (no object)
    \item \textbf{TELL:} He told me/her (that)... (needs object)
    \item Change tenses backwards: present $\rightarrow$ past, past $\rightarrow$ past perfect
    \item Change pronouns, time, and place expressions
    \item \textbf{Questions:} No question word order in reported speech
    \item \textbf{Commands:} told/asked + object + to + infinitive
    \item Don't backshift if it's still true or a general fact
\end{itemize}
